### 2019-10-25 notebooks/20191025_mobility_to_sentence.rmd

---
title: "Mobility Transition Processing"
author: "Dakota Murray"
output: html_notebook
---

```{r}
library(tidyverse)

# Read the transitions file, this will take a few seconds
mobility <- readr::read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Raw/20191024_mobility_transitions.txt", delim = "\t")
```

The total number of publications represented
```{r}
dim(mobility)[1]
```

The head of the file, showing the features
```{r}
head(mobility)
```

Total number and proportion of publications by year
```{r}
t <- table(mobility$pub_year)
t
prop.table(t)
```

Select only the 6 most recent years
```{r}
mobility.2014to2019 <- mobility %>% filter(pub_year >= 2014)
```

Now, lets do the work of creating the mobility sentences
```{r}
mobility2014to2019.processed <- mobility.2014to2019 %>%
  group_by(cluster_id) %>%
  summarize(
    inst_sentence = paste(cwts_org_no, collapse = " ")
  )
```

Top rows showing these mobility sentences
```{r}
head(mobility2014to2019.processed)
```

The total number of mobility sentences
```{r}
dim(mobility2014to2019.processed)[1]
```

Save the result
```{r}
readr::write_csv(mobility2014to2019.processed, path = "/Users/dakotamurray/Dropbox/SME-dropbox/Data/Derived/mobility_sentences/institution_sentences_2019to2019.csv")
```