### 2019-10-25 notebooks/20191025_visualizing_embeddings.rmd

---
title: "Visualizing Institution Mobility Vectors"
author: "Dakota Murray"
output: html_notebook
---

```{r}
library(tidyverse)
library(readr)

inst_coords <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Derived/dim_reduced/inst_umap.csv")
colnames(inst_coords) <- c("rownum", "axis1", "axis2", "token")
inst_lookup <- read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/20191024_institution_lookup.txt", delim = "\t")

country_lookup <- read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/iso_to_country.txt", delim = "\t")
country_lookup
```

```{r}
inst_coords_details <- inst_coords %>%
  left_join(inst_lookup, by = c("token" = "cwts_org_no")) %>%
  left_join(country_lookup, by = c("country_iso_alpha" = "Alpha_code_3"))

inst_coords
inst_coords_details
```

```{r}
org_types <- unique(inst_coords_details$org_type)
org_types_simplified <- c("Univ", "Univ", "Institute", "Institute", "Other", 
                          "Other", "Other", "Other","Hospital", "Other", 
                          "Hospital", "Hospital", "Univ", "Industry")
library(plotly)
g <- inst_coords_details %>%
  filter(!is.na(Continent_name)) %>%
  mutate(
    org = plyr::mapvalues(org_type, org_types, org_types_simplified),
    org = factor(org, levels = c("Univ", "Hospital", "Industry", "Gov", "Other")),
    to_highlight = ifelse(Country_name == "Italy, Italian Republic", T, F)
  ) %>%
  ggplot(aes(x = axis1, y = axis2, color = org, shape = org, group = city, label = full_name, label2 = Country_name, label3 = org_type)) +
  geom_point(size = 0.6, alpha = 0.5, shape = 16) +
  theme_minimal() +
  guides(color = guide_legend(override.aes = list(size=5)),
         shape = guide_legend(override.aes = list(size=5))) +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

ggsave("~/Desktop/mobility_institution_byorg.pdf", g)

ggplotly(g, tooltip = c("Continent_name", "Country_name", "city", "full_name", "org_type"))
```

```{r}
inst_coords <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Derived/dim_reduced/inst_pca.csv")
colnames(inst_coords) <- c("ignore", "axis1", "axis2", "token")
inst_coords_details <- inst_coords %>%
  left_join(inst_lookup, by = c("token" = "cwts_org_no")) %>%
  left_join(country_lookup, by = c("country_iso_alpha" = "Alpha_code_3"))

g <- inst_coords_details %>%
  filter(!is.na(Continent_name)) %>%
  ggplot(aes(x = axis1, y = axis2, color = Continent_name, group = city, label = full_name, label2 = Country_name, label3 = org_type)) +
  geom_point(size = 0.5, alpha = 0.5) +
  theme_minimal() +
  guides(colour = guide_legend(override.aes = list(size=5))) +
  scale_color_brewer(palette = "Set2")

g
ggplotly(g, tooltip = c("Continent_name", "Country_name", "city", "full_name", "org_type"))
```