### 2019-10-28 notebooks/20191025_visualizing_embeddings.rmd
Updated the visualization notebook with new country and city-level visualizations

---
title: "Visualizing Institution Mobility Vectors"
author: "Dakota Murray"
output: html_notebook
---

```{r}
library(tidyverse)
library(readr)

inst_coords <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Figures/UMAP/data/org_2014-2019_umap_data.csv")

inst_lookup <- read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/20191024_institution_lookup.txt", delim = "\t") 
country_lookup <- read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/iso_to_country.txt", delim = "\t")
country_lookup
```

```{r}
inst_coords_details <- inst_coords %>%
  left_join(inst_lookup, by = c("token" = "cwts_org_no")) %>%
  left_join(country_lookup, by = c("country_iso_alpha" = "Alpha_code_3"))

inst_coords
inst_coords_details
```

```{r}
org_types <- unique(inst_coords_details$org_type)
org_types_simplified <- c("Univ", "Univ", "Institute", "Institute", "Other", 
                          "Other", "Other", "Other","Hospital", "Other", 
                          "Hospital", "Hospital", "Univ", "Industry")
library(plotly)
g <- inst_coords_details %>%
  filter(!is.na(Continent_name)) %>%
  mutate(
    org = plyr::mapvalues(org_type, org_types, org_types_simplified),
    org = factor(org, levels = c("Univ", "Hospital", "Industry", "Institute", "Other")),
  ) %>%
  ggplot(aes(x = axis1, y = axis2, color = Continent_name, shape = Continent_name, group = city, label = full_name, label2 = Country_name, label3 = org_type)) +
  geom_point(size = 0.6, alpha = 0.8) +
  theme_minimal() +
  guides(color = guide_legend(override.aes = list(size=5)),
         shape = guide_legend(override.aes = list(size=5))) +
  scale_color_brewer(palette = "Set2") +
  theme(legend.position = "bottom")

#ggsave("~/Desktop/mobility_institution_bycontinent.pdf", g, width = 10, height = 8)

ggplotly(g, tooltip = c("Continent_name", "Country_name", "city", "full_name", "org_type"))
```

```{r}
city_lookup <- inst_lookup %>%
  select(c(city_country, city, country_iso_alpha)) %>%
  group_by(city_country) %>%
  slice(1) %>% 
  rowwise() %>%
  mutate(city_country = gsub("[ ]", "_", city_country))

city_lookup
city_coords <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Figures/UMAP/data/city_2014-2019_umap_data.csv") %>%
  left_join(city_lookup, by = c("token" = "city_country")) %>%
  left_join(country_lookup, by = c("country_iso_alpha" = "Alpha_code_3"))

g <- city_coords %>%
  ggplot(aes(x = axis1, y = axis2,  label = token, color = Continent_name, label2 = Country_name)) +
  geom_point(size = 0.8, alpha = 0.8) +
  theme_minimal() +
  guides(color = guide_legend(override.aes = list(size=5)),
         shape = guide_legend(override.aes = list(size=5))) +
  scale_color_brewer(palette = "Set2") +
  theme(legend.position = "bottom")

ggplotly(g, tooltip = c("token", "name", "Continent_name", "Country_name"))

unique(city_coords$labels)
```

```{r}
country_coords <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Figures/UMAP/data/country_2008-2013_umap_data.csv") %>%
  left_join(country_lookup, by = c("token" = "Alpha_code_3"))

g <- country_coords %>%
  filter(!is.na(Continent_name)) %>%
  ggplot(aes(x = axis1, y = axis2,  label = token, label2 = Country_name, color = Continent_name, shape = Continent_name)) +
  #geom_point(size = 1.2, alpha = 0.8) +
  theme_minimal() +
  geom_text() +
  guides(color = guide_legend(override.aes = list(size=5)),
         shape = guide_legend(override.aes = list(size=5))) +
  scale_color_brewer(palette = "Set2") +
  theme(legend.position = "bottom")

ggplotly(g, tooltip = c("labels", "Continent_name", "Country_name"))
```

```{r}
country_coords <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Figures/UMAP/data/country_2014-2019_umap_data.csv") %>%
  left_join(country_lookup, by = c("token" = "Alpha_code_3"))

g <- country_coords %>%
  filter(!is.na(Continent_name)) %>%
  ggplot(aes(x = axis1, y = axis2,  label = token, label2 = Country_name, color = Continent_name, shape = Continent_name)) +
  #geom_point(size = 1.2, alpha = 0.8) +
  theme_minimal() +
  geom_text() +
  guides(color = guide_legend(override.aes = list(size=5)),
         shape = guide_legend(override.aes = list(size=5))) +
  scale_color_brewer(palette = "Set2") +
  theme(legend.position = "bottom")

ggplotly(g, tooltip = c("labels", "Continent_name", "Country_name"))
```