### 2019-11-05 notebooks/20191104_institution_size_flows.rmd
Notebook experimentin with analysis and visualization

---
title: "Instituion sizes and flows"
author: "Dakota Murray"
output: html_notebook
---

```{r}
library(readr)
library(tidyverse)
sizes <- read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Raw/2008-2019_inst_sizes.txt", delim = "\t")
flows <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Descriptive/org/organization_flows.csv")
lookup <- read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/20191024_institution_lookup.txt", delim = "\t")

geographic_distance <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Descriptive/org/organization_geographic_distance.csv")
word2vec_similarities <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Descriptive/org/2014-2019_word2vec_d200_sim.csv")
```

```{r}
flows <- flows %>%
  select(-index) %>%
  filter(org1 != org2) %>%
  mutate(
    count = ifelse(count == 0, 1, count)
  )

head(flows)
```

```{r}
sizes <- sizes %>%
  group_by(cwts_org_no) %>%
  summarize(
    all_person_count = mean(person_count)
  ) %>%
  arrange(desc(all_person_count))

head(sizes)
```

```{r}
flows_all <- flows %>%
  left_join(sizes, by = c("org1" = "cwts_org_no")) %>%
  left_join(sizes, by = c("org2" = "cwts_org_no")) %>%
  rename(org1_size = all_person_count.x, org2_size = all_person_count.y) %>%
  mutate(gravity = (org1_size * org2_size) / count) %>%
  inner_join(geographic_distance) %>%
  inner_join(word2vec_similarities)

head(flows_all, 100)
```

```{r}
p1_all <- flows_all %>%
  ggplot(aes(x = log(distance), y = log(gravity))) +
  geom_hex(bins = 200) +
  theme_minimal() +
  labs(x = "Log(Kilometers distance)",
       y = "P1 * P2 / F12")

p1_all
```

```{r}
p2_all <- flows_all %>%
  ggplot(aes(x = (similarity), y = log(gravity))) +
  geom_hex(bins = 100) +
  theme_minimal() +
  labs(x = "Cosine Similarity",
       y = "Log(Kilometer Distance)")

p2_all
```

```{r}
p1_

```

```{r}
library(grid)
library(gridExtra)

```