### 2020-01-08 notebooks/20191104_institution_size_flows.rmd

---
title: "Instituion sizes and flows"
author: "Dakota Murray"
output: html_notebook
---

```{r}
library(readr)
library(tidyverse)
sizes <- read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Raw/2008-2019_inst_sizes.txt", delim = "\t")
flows <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Descriptive/org/organization_flows.csv")
lookup <- read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/20191024_institution_lookup.txt", delim = "\t")

geographic_distance <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Descriptive/org/organization_geographic_distance.csv")
word2vec_similarities <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Descriptive/org/2014-2019_word2vec_d200_sim.csv")
```

```{r}
flows <- flows %>%
  filter(org1 != org2) %>%
  mutate(
    count = ifelse(count == 0, 1, count)
  )

head(flows)
```

```{r}

lookup <- read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/20191024_institution_lookup.txt",
                     delim = "\t")

coords <- read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/fixed_org_coordinates.txt", delim = "\t")

lookup %>%
  select(-new_lat, -new_long) %>%
  left_join(coords, by = "cwts_org_no") %>%
  mutate(
    latitude = ifelse(latitude == "NULL", new_lat, latitude),
    longitude = ifelse(longitude == "NULL", new_long, longitude),
    ) %>%
  select(-new_lat, -new_long)

```

```{r}
sizes <- sizes %>%
  group_by(cwts_org_no) %>%
  summarize(
    all_person_count = mean(person_count)
  ) %>%
  arrange(desc(all_person_count))

head(sizes)
```

```{r}

lookup <- read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/institution_lookup_with_states.txt", delim = "\t")

lookup_trim <- lookup %>% select(cwts_org_no, city, region, country_iso_name)
flows_all <- flows %>%
  left_join(sizes, by = c("org1" = "cwts_org_no")) %>%
  left_join(sizes, by = c("org2" = "cwts_org_no")) %>%
  rename(org1_size = all_person_count.x, org2_size = all_person_count.y) %>%
  mutate(gravity = (org1_size * org2_size) / count) %>%
  inner_join(geographic_distance) %>%
  inner_join(word2vec_similarities) %>%
  left_join(lookup_trim, by = c("org1" = "cwts_org_no")) %>%
  rename(org1_city = city, org1_region = region, org1_country = country_iso_name) %>%
  left_join(lookup_trim, by = c("org2" = "cwts_org_no")) %>%
  rename(org2_city = city, org2_region = region, org2_country = country_iso_name)

head(flows_all, 100)
```

```{r}
library(tidyverse)
library(readr)
library(latex2exp)
library(ggpmisc)

lookup <- read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/institution_lookup_with_states.txt", delim = "\t")
lookup_trim <- lookup %>% select(cwts_org_no, city, region, country_iso_name)

dist_raw = read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Descriptive/org/aggregate_org_distances.csv") %>%
  select(-X1) %>%
  left_join(lookup_trim, by = c("org1" = "cwts_org_no")) %>%
  rename(org1_city = city, org1_region = region, org1_country = country_iso_name) %>%
  left_join(lookup_trim, by = c("org2" = "cwts_org_no")) %>%
  rename(org2_city = city, org2_region = region, org2_country = country_iso_name) %>%
  unique()

head(dist_raw) 
```

```{r}
dist <- dist_raw %>% 
  #filter(org1_city != org2_city) %>%
  filter(emb_similarity >= 0) %>%
  filter(count > 0) %>%
  mutate(
         geo_distance = ifelse(geo_distance < 1, 1, geo_distance),
         geo_distance_logged = log(geo_distance),
         gravity = imputed_count / (org1_size * org2_size),
         gravity_logged = log(gravity)
     )

dist_binned <- dist %>%
  gather(metric, distance, geo_distance_logged, emb_similarity) %>%
  mutate(metric = factor(metric, 
                         levels = c("geo_distance_logged", "emb_similarity"), 
                         labels = c("log(km distance)", "cosine similarity"))
  ) %>%
  group_by(metric) %>%
  mutate(
    bin = cut(round(distance, 2), 40)
  ) %>%
  arrange(bin) %>%
  group_by(metric, bin) %>%
  summarize(
    pos = min(distance) + ((max(distance) - min(distance)) / 2),
    #pos = (as.numeric(first(bin)) * 0.05) - 0.025,
    mu = mean(gravity_logged, na.rm = T),
    se = sd(gravity_logged, na.rm = T) / sqrt(n()),
    ci = 2.576 * se
  )

dist_binned

dist %>%
  gather(metric, distance, geo_distance_logged, emb_similarity) %>%
  mutate(metric = factor(metric, 
                         levels = c("geo_distance_logged", "emb_similarity"), 
                         labels = c("log(km distance)", "cosine similarity"))
  ) %>%
  ggplot(aes(x = distance, y = gravity_logged)) +
  geom_hex(bins = 20, 
           color = "lightgrey") +
  stat_poly_eq(formula = y ~ x,
               aes(label = paste(..rr.label.., sep = "~~~")), 
               parse=TRUE,
               coef.digits = 2,
               label.x.npc = "right") +
  stat_smooth(method = "lm", formula = y ~ x, color = "#c0392b", size = 1.5) + 
  geom_point(data = dist_binned, aes(x = pos, y = mu, group = bin), size = 2) +
  geom_errorbar(data = dist_binned, aes(x = pos, ymin = mu - ci, ymax = mu + ci, y = NULL)) +
  facet_wrap(~metric, scale = "free_x", strip.position = "bottom") +
  theme_minimal() +
  scale_fill_gradientn(colours=c("white", "blue"), 
                       name = "Frequency",
                       na.value=NA) +
  theme(
    legend.position = "right",
    strip.text = element_text(size = 12),
    axis.title.y = element_text(size = 12, face = "bold", angle = 0, vjust = 0.5),
    axis.title.x = element_blank()
  ) +
  ylab(TeX("$\\log\\left(\\frac{F_{ij}}{P_{i}P_{j}}\\right)$"))
```

```{r}
g1 <- dist %>% 
  filter(count > 0) %>%
  mutate(
         geo_distance = ifelse(geo_distance < 1, 1, geo_distance),
         geo_distance_logged = log(geo_distance),
         gravity = imputed_count / (org1_size * org2_size),
         gravity_logged = log(gravity)
         ) %>%
  gather(metric, distance, geo_distance_logged, emb_similarity) %>%
  mutate(metric = factor(metric, 
                         levels = c("geo_distance_logged", "emb_similarity"), 
                         labels = c("log(km distance)", "cosine similarity"))
  ) %>%
  ggplot(aes(x = gravity_logged, y = distance)) +
  geom_hex(bins = 20, 
           color = "lightgrey") +
  stat_poly_eq(formula = y ~ x,
               aes(label = paste(..rr.label.., sep = "~~~")), 
               parse=TRUE,
               coef.digits = 2,
               label.x.npc = "right") +
  stat_smooth(method = "lm", formula = y ~ x, color = "#c0392b", size = 1.5) + 
  facet_wrap(~metric, scale = "free_y", strip.position = "left") +
  theme_minimal() +
  scale_fill_gradientn(colours=c("white", "blue"), 
                       name = "Frequency",
                       na.value=NA) +
  theme(
    legend.position = "right",
    strip.text = element_text(size = 12),
    axis.title.x = element_text(size = 12, face = "bold"),
    axis.title.y = element_blank()
  ) +
  xlab(TeX("$\\log\\left(\\frac{F_{ij}}{P_{i}P_{j}}\\right)$")) +
  ylab("Metric")

g1
ggsave("~/Desktop/plot1.png", g1, width = 9, height = 5)
```

We can also look at log ratios. Note that I filter to only values of cosine similarity that are at least 0. The log function is only defined for non-negative numbers, however cosine simialrity is defined between -1 and 1 (though there are relatively few negative comparisons).

Higher values on the y-axis correpond to cases where cosine similarity is much greater than the gravity, i.e., the cosine similarity indicates stronger relationship than the flow.

The "gravity" is the flow, as a factor of the "mass" or size of the institions. Higher values suggest that the flow is greater than expected given the institution sizes. 
```{r}
dist %>%
  filter(emb_similarity >= 0) %>%
  mutate(
    gravity = imputed_count / (org1_size * org2_size),
    cosine_bin = cut(round(emb_similarity, 2), 20),
    log_ratio = log(emb_similarity / gravity)
  ) %>%
  arrange(cosine_bin) %>%
  group_by(cosine_bin) %>%
  summarize(
    mu_lr = mean(log_ratio, na.rm = T),
    se_lr = sd(log_ratio, na.rm = T) / sqrt(n())
  ) %>%
  ggplot(aes(x = cosine_bin, y = mu_lr)) +
  geom_point(size = 4, alpha = 0.8, shape = 21) +
  geom_errorbar(aes(ymin = mu_lr - se_lr, ymax = mu_lr + se_lr), width = 0.4) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
    axis.title.y = element_text(angle = 0, vjust = 0.5)
  ) +
  xlab("Cosine similarity (binned)") +
  ylab(TeX("$\\log\\left[\\frac{\\cos(\\theta)}{\\frac{F_{ij}}{P_{i}P_{j}}}\\right]"))
```

```{r}
dist2 <- dist %>%
  filter(emb_similarity >= 0) %>%
  mutate(
    sim = 1 - emb_similarity,
    y = sim, 
    x = imputed_count / (org1_size * org2_size),
  )

fitted <- nls(y ~ a * exp(b * x),
    data = dist2,
    start=list(a=0.5, b=-0.1))

```

```{r}
dist %>%
  filter(count > 0) %>%
  filter(emb_similarity >= 0) %>%
  mutate(
    gravity = imputed_count / (org1_size * org2_size),
    sim = 1 - emb_similarity
  ) %>%
  ggplot(aes(x = gravity, y = sim)) +
  geom_hex(bins = 30) +
  geom_smooth(method = "nls", formula = y ~ a * exp(b * x), method.args = list(start=c(a = 0.5, b= -0.01)), se = FALSE) +
  theme_minimal() +
  labs(x = "f12 / (p1 * p2), not transformed", y = "(1 - cosine similarity)")
```

```{r}
dist %>%
  filter(emb_similarity >= 0) %>%
  ggplot(aes(x = geo_distance_logged, y = gravity_logged)) +
  geom_hex(bins = 20, 
           color = "lightgrey") +
  #geom_point(alpha = 0.1) +
  stat_poly_eq(formula = y ~ x,
               aes(label = paste(..rr.label.., sep = "~~~")), 
               parse=TRUE,
               coef.digits = 2,
               label.x.npc = "right") +
  stat_smooth(method = "lm", formula = y ~ x, color = "#c0392b", size = 1.5) + 
  theme_minimal() +
  scale_fill_gradientn(colours=c("white", "blue"), 
                       name = "Frequency",
                       na.value=NA) +
  theme(
    legend.position = "right",
    strip.text = element_text(size = 12),
    axis.title.x = element_text(size = 12, face = "bold"),
    axis.title.y = element_text(size = 12, face = "bold", angle = 0, vjust = 0.5)
  ) +
  xlab("Log(Geo. Distance, km)") +
  ylab(TeX("$\\log\\left(\\frac{F_{ij}}{P_{i}P_{j}}\\right)$"))

```

```{r}
lookup_country <- lookup %>%
  select(cwts_org_no, country_iso_name, city_country)
dist_geo <- dist %>%
  left_join(lookup_country, by = c("org1" = "cwts_org_no")) %>%
  rename(country1 = country_iso_name, city1 = city_country) %>%
  left_join(lookup_country, by = c("org2" = "cwts_org_no")) %>%
  rename(country2 = country_iso_name, city2 = city_country) %>%
  mutate(
    same_country = country1 == country2,
    same_city = city1 == city2
  ) %>%
  filter()

dist_geo
```

```{r}
completed <- c(10410, 1292, 18381, 3505, 18538, 18480, 15219, 846, 1742, 1868, 3047, 15229, 15202, 18503, 15929, 18178, 15980, 15939, 3962, 15852, 15865, 3961, 15966, 3961, 1247, 10384, 3022, 30218, 30219, 818, 2007, 3109, 1062, 10345, 1076, 3614, 7377, 427, 31919, 31913, 38, 2625, 10358, 2628, 1846, 1847, 2732, 2663, 3112, 2662, 2738, 31306)
dist_geo %>%
  filter(same_country == T) %>%
  filter(!org1 %in% completed) %>%
  filter(!org2 %in% completed) %>%
  group_by(country1) %>%
  mutate(
    mu = mean(geo_distance),
    #top_percentile = quantile(geo_distance, seq(0,1, by=0.01))[100],
    #diff = abs(mu - geo_distance),
    #is_anomoly = geo_distance > (top_percentile)
    is_anomoly = geo_distance > ( 15 * mu)
  ) %>%
  filter(is_anomoly)

```

```{r}
lookup %>%
  group_by(country_iso_name) %>%
  mutate(
    mean_lat = mean(as.numeric(latitude), na.rm = T),
    mean_long = mean(as.numeric(longitude), na.rm = T),
    is_anomoly = abs(latitude - mean_lat) > 10 | abs(longitude - mean_long) > 10
  ) %>%
  filter(is_anomoly == T)

```