### 2020-02-18 notebooks/20200207_plotting_notebook.rmd

---
title: "Model Hyperparameters"
output: html_notebook
---

```{r}
library(tidyverse)
library(readr)

sims <- read_csv("/Users/dakotamurray/Desktop/sims_coasts.csv")

lookup <- read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/data/Additional/institution_lookup_with_states.txt", delim = "\t", col_types = readr::cols())

sims <- sims %>%
  left_join(lookup, by = "cwts_org_no") %>%
  distinct(cwts_org_no, .keep_all = T)

coasts <- sims
```

```{r fig.width=3, fig.height=1}
averages <- sims %>%
  group_by(region) %>%
  summarize(
    mu = mean(sim)
  ) %>%
  ungroup() %>%
  filter(region %in% c("Arizona", "Connecticut"))

sims %>%
  mutate(
    highlight = ifelse(region == "Connecticut", "CT", 
                       ifelse(region == "Arizona", "AZ", "other"))
  ) %>%
  ggplot(aes(x = sim, y = 0, color = highlight, alpha = ifelse(highlight == "other", 0.9, 1))) +
  geom_segment(y = -1, yend = 1, aes(xend = sim)) +
  geom_hline(yintercept = 0, size = 1) +
  geom_segment(y = -0.2, yend = 0.2, x = 0, xend = 0, size = 1, color = "black") +
  geom_point(data = averages, 
             aes(x = mu, y = 0),
             color = ifelse(averages$region == "Arizona", "#c0392b", "#2980b9"),
             alpha = 1,
             size = 5,
             shape = 21,
             stroke = 2
  ) +
  xlim(-0.3, 0.3) +
  scale_y_continuous(
    limits = c(-1.1, 1.1), 
    name = "   California",
    sec.axis = dup_axis(name = "Massachusetts")
  ) +
  scale_color_manual(values = c("#c0392b", "#2980b9", "grey")) +
  guides(alpha = F,
         color = guide_legend(override.aes = list(size = 2))
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    axis.text.y = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 11),
    axis.title.y.left = element_text(angle = 0, vjust = 0.5),
    axis.title.y.right = element_text(angle = 0, vjust = 0.5),
    panel.grid = element_blank()
  )
```

```{r}
sims <- read_csv("/Users/dakotamurray/Desktop/sims_statevuc.csv")

sims <- sims %>%
  left_join(lookup, by = "cwts_org_no") %>%
  distinct(cwts_org_no, .keep_all = T)
```

```{r fig.width=3, fig.height=1}
elite <- c(1277, 1187, 1255, 1188, 1251, 1296, 1193, 1272, 1192, 1208)
nonelite <- c(3501, 18086, 18530, 18486, 1471, 1468, 1558, 1287, 3538, 18373, 9856, 3482, 1268, 10379, 18514)
averages <- sims %>%
  mutate(
    elite = ifelse(cwts_org_no %in% elite, "Elite", 
                   ifelse(cwts_org_no %in% nonelite, "Not elite", "other")
            )
    ) %>%
  filter(elite != "other") %>%
  group_by(elite) %>%
  summarize(
    mu = mean(sim)
  ) %>%
  ungroup()

sims %>%
  mutate(
    elite = ifelse(cwts_org_no %in% elite, "Elite", ifelse(cwts_org_no %in% nonelite, "Not elite", "other"))
  ) %>%
  ggplot(aes(x = sim, y = 0, color = elite, alpha = ifelse(elite %in% c("Elite", "Not elite"), 1, 0.9))) +
  geom_segment(y = -1, yend = 1, aes(xend = sim)) +
  geom_hline(yintercept = 0, size = 1) +
  geom_segment(y = -0.2, yend = 0.2, x = 0, xend = 0, size = 1, color = "black") +
  geom_point(data = averages, 
             aes(x = mu, y = 0),
             color = ifelse(averages$elite == "Elite", "#f39c12", "#27ae60"),
             alpha = 1,
             size = 5,
             shape = 21,
             stroke = 2
  ) +
  xlim(-0.3, 0.3) +
  scale_y_continuous(
    limits = c(-1.1, 1.1), 
    name = "Cal. State",
    sec.axis = dup_axis(name = "Univ. Cal.")
  ) +
  scale_color_manual(values = c("#f39c12", "#27ae60", "grey")) +
  guides(alpha = F,
         color = guide_legend(override.aes = list(size = 2))
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    axis.text.y = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 11),
    axis.title.y.left = element_text(angle = 0, vjust = 0.5),
    axis.title.y.right = element_text(angle = 0, vjust = 0.5),
    panel.grid = element_blank()
  )
```

```{r}
sims <- read_csv("/Users/dakotamurray/Desktop/sims_prestige.csv")

sims <- sims %>%
  left_join(lookup, by = "cwts_org_no") %>%
  distinct(cwts_org_no, .keep_all = T)

prestige <- sims
```

```{r fig.width=3, fig.height=1}
averages <- sims %>%
  group_by(region) %>%
  summarize(
    mu = mean(sim)
  ) %>%
  ungroup() %>%
  filter(region %in% c("Indiana", "Maryland"))

sims %>%
  mutate(
    highlight = ifelse(region %in% c("Indiana", "Maryland"), region, "other")
  ) %>%
  ggplot(aes(x = sim, y = 0, color = highlight, alpha = ifelse(highlight == "other", 0.9, 1))) +
  geom_segment(y = -1, yend = 1, aes(xend = sim)) +
  geom_hline(yintercept = 0, size = 1) +
  geom_segment(y = -0.2, yend = 0.2, x = 0, xend = 0, size = 1, color = "black") +
  geom_point(data = averages, 
             aes(x = mu, y = 0),
             color = ifelse(averages$region == "Indiana", "#8e44ad", "#22a6b3"),
             alpha = 1,
             size = 5,
             shape = 21,
             stroke = 2
  ) +
  xlim(-0.3, 0.3) +
  scale_y_continuous(
    limits = c(-1.1, 1.1), 
    name = "Non Elite",
    sec.axis = dup_axis(name = "Elite   ")
  ) +
  scale_color_manual(values = c("#8e44ad", "#22a6b3", "grey")) +
  guides(alpha = F,
         color = guide_legend(override.aes = list(size = 2))
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    axis.text.y = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 11),
    axis.title.y.left = element_text(angle = 0, vjust = 0.5),
    axis.title.y.right = element_text(angle = 0, vjust = 0.5),
    panel.grid = element_blank()
  )
```

```{r fig.width=5, fig.height=3}
library(plotly)

all <- prestige %>%
  select(cwts_org_no, sim) %>%
  left_join(coasts, by = "cwts_org_no") %>%
  rename(sim.prestige = sim.x, sim.coasts = sim.y)

to_color <- c("Indiana", "Ohio", "Connecticut", "New York", "Arizona", "California", "Massachusetts", "Florida")

labels <- readr::read_csv("/Users/dakotamurray/Desktop/axis_unis_labels.csv") %>%
  inner_join(all, by = c("cwts_org_no"))

p <- all %>%
  filter(org_type_code == "U") %>%
  mutate(
    state = ifelse(region %in% to_color, region, "Others"),
    state = factor(state, levels = c(sort(to_color), "Others"))
  ) %>%
  filter(state != "Others") %>%
  ggplot(aes(x = sim.coasts, y = sim.prestige, fill = state, label1 = full_name, label2 = region, label3 = city)) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(size = 3.5, shape = 21) +
  ggrepel::geom_label_repel(data = labels, 
             inherit.aes = F,
             aes(x = sim.coasts, y = sim.prestige, label = short_name)
             ) +
  scale_fill_manual(name = "Region", values = c("#7fc97f", "#beaed4", "#fdc086", "#ffff99", "#386cb0", "#f0027f", "#bf5b17", "#666666")) +
  scale_x_continuous(
    limits = c(-0.4, 0.4),
    name = "Lower Research Impact",
    sec.axis = dup_axis(name = "High Research Impact")
  ) +
  scale_y_continuous(
    limits = c(-0.42, 0.42),
    name = "California",
    sec.axis = dup_axis(name = "Massachusetts")
  ) +
  theme_minimal() +
  theme(
    text = element_text(family = "Helvetica", size = 11),
    axis.title.x = element_text(angle = 0, size = 12, face = "bold", vjust = 0.5),
    axis.title.x.top = element_text(angle = 0, size = 12, face = "bold", vjust = 0.5),
    axis.title.y = element_text(angle = 0, size = 12, face = "bold", vjust = 0.5),
    axis.title.y.right = element_text(angle = 0, size = 12, face = "bold", vjust = 0.5),
    legend.text = element_text(size = 12),
    legend.title = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom"
  )

p
widget <- ggplotly(p, tooltip = c("full_name", "region", "city"))
widget
htmlwidgets::saveWidget(as.widget(widget), "~/Desktop/UniCoastPrestigeAxis.html")
```