### 2020-02-20 notebooks/20190116_notebook_descriptive_info.Rmd

---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(readr)

mobile <- read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/data/Derived/Trajectories/precedence_mobility_trajectories.txt", delim = "\t") 
nonmobile <- read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/data/raw/nonmobile_researcher_trajectories.txt", delim = "\t")

dist <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/data/Derived/Descriptive/org/aggregate/precedence/aggregate_org_distances_precedence_d300_ws1.csv")

flows <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/data/Derived/Descriptive/org/precedence_organization_flows.csv")
```

The total number of co-affiliations (mobile and nonmobile)
```{r}
dim(mobile)[1] + dim(nonmobile)[1]
```

The total number of authors
```{r}
length(unique(mobile$cluster_id)) + length(unique(nonmobile$cluster_id))
```

```{r}
mobile_real <- mobile %>%
  group_by(cluster_id) %>%
  filter(length(unique(cwts_org_no)) > 1) %>%
  ungroup()
```

Total number of mobile authors
```{r}
length(unique(mobile_real$cluster_id))
```

Percentage of mobile authors of all authors
```{r}
length(unique(mobile_real$cluster_id)) / (length(unique(mobile$cluster_id)) + length(unique(nonmobile$cluster_id)))
```

Total number of co-affiliation links
```{r}
dim(mobile_real)[1]
```

Number of unique organizations
```{r}
length(unique(c(nonmobile$cwts_org_no, mobile$cwts_org_no)))
```

Number of unique organizations among nonmobile researchers
```{r}
length(unique(nonmobile$cwts_org_no))
```

The longest trajectory among mobile scholars
```{r}
counted <- mobile_real %>%
  group_by(cluster_id) %>%
  summarize(
    count = length(unique(cwts_org_no))
  )

# Mean
mean(counted$count)

# Median 
median(counted$count)

# Max
max(counted$count)

```

Total number of organizations in the final, embedded dataset
```{r}
length(unique(c(dist$org1, unique(dist$org2))))
```

Percentage of total unique organizations covered by the embeddingc
```{r}
length(unique(c(dist$org1, unique(dist$org2)))) / length(unique(c(nonmobile$cwts_org_no, mobile$cwts_org_no)))

```

```{r}

```