### 2020-07-24 notebooks/semaxis_notebook.rmd
updated with table, new supp figure, and new text related to fig 6

---
title: "SemAxis plotting"
output: html_notebook
---

```{r}
org_types <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/org_types.csv")
lookup <- read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/institution_lookup_with_states.txt", delim = "\t", col_types = readr::cols()) %>%
  filter(country_iso_alpha == "USA") %>%
  left_join(org_types, by = c("org_type")) %>%
  filter(org_type_simplified == "Government") %>%
  select(cwts_org_no, org_type, full_name)

lookup %>%
  filter(org_type == "Funding Organisation")
```

```{r}
library(readr)
library(tidyverse)
org_types <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/org_types.csv")
# Load organization meta-info
lookup <- read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/institution_lookup_with_states.txt", delim = "\t", col_types = readr::cols()) %>%
  filter(country_iso_alpha == "USA") %>% # select only the specified country
  select(cwts_org_no, country_iso_alpha, region, org_type, org_type_code, full_name) %>%
  left_join(org_types, by = "org_type")
  #filter(org_type_simplified == "Institute") # select only universities
  #filter(org_type_simplified == "Institute" | org_type_simplified == "Government") # select only universities

axis2 <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Derived/SemAxis/precedence/prestige/times_d300_ws1_n20_precedence_semaxis.csv", col_types = readr::cols())

# Load the axis data and filter to only specified countries
sims <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Derived/SemAxis/precedence/geography/coasts_d300_ws1_n20_semaxis.csv", col_types = readr::cols()) %>%
  inner_join(lookup, by = "cwts_org_no") %>%
  inner_join(axis2, by = "cwts_org_no")

```

```{r}
plotdata <- sims
library(plotly)

plotdata %>%
  filter(org_type_simplified == "Institute") %>%
  top_n(-1, sim.y)

p <- plotdata %>%
  ggplot() +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(aes(x = sim.x, y = sim.y, label1 = full_name, label2 = cwts_org_no),
             size = 3.5,
             shape = 16,
             fill = "lightgrey",
             alpha = 0.2,
             data = subset(plotdata, org_type_simplified == "Government")
           ) +
  #geom_point(aes(x = sim.x, y = sim.y, label = full_name),
  #           size = 3.5,
  #           shape = 21,
  #           fill = "#f39c12",
  #           alpha = 1.0,
  #           data = subset(plotdata, org_type_simplified == "Government")
  #         ) +
  theme_minimal() +
  theme(
    text = element_text(family = "Helvetica", size = 11),
    axis.title.x = element_text(angle = 0, size = 12, face = "bold", vjust = 0.5),
    axis.title.x.top = element_text(angle = 0, size = 12, face = "bold", vjust = 0.5),
    axis.title.y = element_text(angle = 0, size = 12, face = "bold", vjust = 0.5),
    axis.title.y.right = element_text(angle = 0, size = 12, face = "bold", vjust = 0.5),
    legend.text = element_text(size = 12),
    legend.title = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom"
  )

p

widget <- ggplotly(p, tooltip = c("full_name", "cwts_org_no", "region", "city"))
widget
```

```{r}
# Load organization meta-info
lookup <- read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/institution_lookup_with_states.txt", delim = "\t", col_types = readr::cols()) %>%
  filter(country_iso_alpha == "GBR" & org_type_code == "U") %>% # select only the specified country
  select(cwts_org_no, country_iso_alpha, region, org_type, org_type_code, full_name)

axis2 <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Derived/SemAxis/precedence/prestige/leiden_d300_ws1_n20_precedence_semaxis.csv", col_types = readr::cols())

# Load the axis data and filter to only specified countries
sims <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Derived/SemAxis/precedence/geography/coasts_d300_ws1_n20_semaxis.csv", col_types = readr::cols()) %>%
  inner_join(lookup, by = "cwts_org_no") %>%
  inner_join(axis2, by = "cwts_org_no")

```

```{r}
plotdata <- sims
library(plotly)

p <- plotdata %>%
  ggplot() +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(aes(x = sim.x, y = sim.y, label = full_name),
             size = 3.5,
             shape = 16,
             fill = "lightgrey",
             alpha = 0.2,
             #data = subset(plotdata, org_type_simplified == "Institute")
           ) +
  theme_minimal() +
  theme(
    text = element_text(family = "Helvetica", size = 11),
    axis.title.x = element_text(angle = 0, size = 12, face = "bold", vjust = 0.5),
    axis.title.x.top = element_text(angle = 0, size = 12, face = "bold", vjust = 0.5),
    axis.title.y = element_text(angle = 0, size = 12, face = "bold", vjust = 0.5),
    axis.title.y.right = element_text(angle = 0, size = 12, face = "bold", vjust = 0.5),
    legend.text = element_text(size = 12),
    legend.title = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom"
  )

p

widget <- ggplotly(p, tooltip = c("full_name", "region", "city"))
widget
```

```{r}
# Load organization meta-info
lookup <- read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/institution_lookup_with_states.txt", delim = "\t", col_types = readr::cols()) %>%
  filter(country_iso_alpha %in% c("GBR", "JPN") & org_type_code == "U") %>% # select only the specified country
  select(cwts_org_no, country_iso_alpha, region, org_type, org_type_code, full_name)

axis2 <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Derived/SemAxis/precedence/prestige/leiden_d100_ws1_n1_precedence_semaxis.csv", col_types = readr::cols())

# Load the axis data and filter to only specified countries
sims <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Derived/SemAxis/precedence/geography/coasts_d100_ws1_n20_semaxis.csv", col_types = readr::cols()) %>%
  inner_join(lookup, by = "cwts_org_no") %>%
  inner_join(axis2, by = "cwts_org_no")

```

```{r}
plotdata <- sims
library(plotly)

p <- plotdata %>%
  ggplot() +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(aes(x = sim.x, y = sim.y, label = full_name, fill = country_iso_alpha),
             size = 3.5,
             shape = 21,
             #fill = "lightgrey",
             alpha = 0.8,
             #data = subset(plotdata, org_type_simplified == "Institute")
           ) +
  theme_minimal() +
  theme(
    text = element_text(family = "Helvetica", size = 11),
    axis.title.x = element_text(angle = 0, size = 12, face = "bold", vjust = 0.5),
    axis.title.x.top = element_text(angle = 0, size = 12, face = "bold", vjust = 0.5),
    axis.title.y = element_text(angle = 0, size = 12, face = "bold", vjust = 0.5),
    axis.title.y.right = element_text(angle = 0, size = 12, face = "bold", vjust = 0.5),
    legend.text = element_text(size = 12),
    legend.title = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom"
  )

p

widget <- ggplotly(p, tooltip = c("full_name", "region", "city"))
widget
```

```{r}

times <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/times_ranking.csv")
leiden <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/leiden_ranking.csv")

lookups <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/institution_lookup_with_states.txt")

sims <- lookup %>%
  filter(country_iso_alpha == "USA") %>%
  left_join(times, by = c("cwts_org_no")) %>%
  inner_join(leiden, by = c("cwts_org_no")) %>%
  select(full_name, cwts_org_no, total_score, impact_frac_mncs) %>%
  arrange(desc(total_score)) %>%
  mutate(
    times_rank = row_number()
  ) %>%
  arrange(desc(impact_frac_mncs)) %>%
  mutate(
    leiden_rank = row_number()
  )

# Build the plot
plot <- sims %>%
  ggplot(aes(x = times_rank, y = leiden_rank)) +
  geom_abline() +
  geom_point(size = 3.5, stroke = 0.5) +
  # Add a fake top axis title, just to ensure that its the same size as the 2d fig
  scale_x_continuous(limits = c(0, 150), sec.axis = dup_axis(name = "")) +
  scale_y_continuous(limits = c(0, 150)) +
  scale_shape_manual(values = c(1, 16)) +
  guides(shape = F) +
  theme_minimal() +
  theme(
    text = element_text(size = 12, family = "Helvetica"),
    axis.text.x.top = element_blank(),
    axis.title = element_text(size = 14, face = "bold"),
    panel.grid.minor = element_blank(),
  ) +
  ylab("") +
  xlab("SemAxis Rank")

plot

```