### 2021-02-26 notebooks/metric_performance_tables.rmd

---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

```{r}
library(tidyverse)
library(readr)
library(huxtable)
rmse <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Derived/Stat/aggregate_rmse.csv")
```

```{r}
r2 <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Derived/Stat/aggregate_r2.csv")

metric_levels <- c("emb", "levyeuc", "dot", "pprjsd", "svdcos", "levycos", "geo",
                   "lapcos", "pprcos", "levydot")

metric_labels <- c("Embedding cosine", "Factorized euclidean", "Embedding dot", "PPR JSD", 
                   "SVD cosine", "Factorized cosine", "Geographic", "Laplacian cosine", "PPR cosine", "Factorized dot")

r2.tab <- r2 %>%
  filter(traj == "precedence") %>%
  filter(dim == 300) %>%
  filter(ws == 1) %>%
  filter(gamma == 1.0) %>%
  select(metric, case, sizetype, r2) %>%
  arrange(desc(r2)) %>%
  mutate(
    case = factor(case, levels = c("global", "same-country", "diff-country")),
    sizetype = factor(sizetype, levels = c("all", "mobile", "freq"))
  ) %>%
  arrange(case, sizetype, desc(r2)) %>%
  group_by(case, sizetype) %>%
  mutate(
    r2 = ifelse(max(r2) == r2, paste0("*", round(r2, 3)), as.character(round(r2, 3)))
  ) %>%
  group_by(metric, case, sizetype) %>%
  ungroup() %>%
  pivot_wider(names_from = c("case", "sizetype"), values_from = "r2") %>%
  #select(-id) %>%
  mutate(
    metric = factor(metric, metric_levels, metric_labels)
  ) %>%
  rename(" " = metric)

r2.hux <- as_hux(r2.tab) %>%
  set_contents(1, 2:10, rep(c("Global", "Domestic", "International"), 3)) %>%
  insert_row("", "All", "", "", "Mobile only", "", "", "Raw freq", "", "", after = 0) %>% 
  merge_cells(1, 2:4) %>% 
  merge_cells(1, 5:7) %>% 
  merge_cells(1, 8:10) %>%
  set_tb_padding(1, everywhere, 0) %>% 
  set_bold(1, everywhere) %>%
  map_align(by_cols(c("left", rep("right", 9)))) %>%
  set_align(1, everywhere, "center") %>% 
  #map_align(row = 1, by_cols(rep("center", 10))) %>%
  set_header_rows(1:2, TRUE) %>% 
  set_header_cols(1, TRUE) %>% 
  style_headers(bold = TRUE) %>%
  set_right_border(1:12, 1, value = 0.5) %>%
  set_right_border(1:12, 4, value = 0.5) %>%
  set_right_border(1:12, 7, value = 0.5) %>%
  set_bottom_border(2, 1:10, value = 0.5)

r2.hux %>%
  print_latex()
```

```{r}
rmse <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Derived/Stat/aggregate_rmse.csv")

metric_levels <- c("emb", "levyeuc", "dot", "pprjsd", "svdcos", "levycos", "geo",
                   "lapcos", "pprcos", "levydot")

metric_labels <- c("Embedding cosine", "Factorized euclidean", "Embedding dot", "PPR JSD", 
                   "SVD cosine", "Factorized cosine", "Geographic", "Laplacian cosine", "PPR cosine", "Factorized dot")

rmse.tab <- rmse %>%
  filter(dim == 300) %>%
  filter(ws == 1) %>%
  filter(gamma == 1.0) %>%
  rename(rmse = error.exp) %>%
  #gather(model, rmse, error.exp, error.power) %>%
  select(metric, case, sizetype, rmse) %>%
  arrange(rmse) %>%
  mutate(
    case = factor(case, levels = c("global", "same", "different")),
    sizetype = factor(sizetype, levels = c("all", "mobile", "freq"))
  ) %>%
  arrange(case, sizetype, rmse) %>%
  group_by(case, sizetype) %>%
  mutate(
    rmse = ifelse(min(rmse) == rmse, paste0("*", round(rmse, 3)), as.character(round(rmse, 3)))
  ) %>%
  group_by(metric, case, sizetype) %>%
  ungroup() %>%
  pivot_wider(names_from = c("case", "sizetype"), values_from = "rmse") %>%
  #select(-id) %>%
  mutate(
    metric = factor(metric, metric_levels, metric_labels)
  ) %>%
  rename(" " = metric)

rmse.hux <- as_hux(rmse.tab) %>%
  set_contents(1, 2:10, rep(c("Global", "Domestic", "International"), 3)) %>%
  insert_row("", "All", "", "", "Mobile only", "", "", "Raw freq", "", "", after = 0) %>% 
  merge_cells(1, 2:4) %>% 
  merge_cells(1, 5:7) %>% 
  merge_cells(1, 8:10) %>%
  set_tb_padding(1, everywhere, 0) %>% 
  set_bold(1, everywhere) %>%
  map_align(by_cols(c("left", rep("right", 9)))) %>%
  set_align(1, everywhere, "center") %>% 
  #map_align(row = 1, by_cols(rep("center", 10))) %>%
  set_header_rows(1:2, TRUE) %>% 
  set_header_cols(1, TRUE) %>% 
  style_headers(bold = TRUE) %>%
  set_right_border(1:12, 1, value = 0.5) %>%
  set_right_border(1:12, 4, value = 0.5) %>%
  set_right_border(1:12, 7, value = 0.5) %>%
  set_bottom_border(2, 1:10, value = 0.5)

rmse.hux %>%
  print_latex()

```

Now the same table, with the power-law model
```{r}
rmse.tab <- rmse %>%
  filter(dim == 300) %>%
  filter(ws == 1) %>%
  filter(gamma == 1.0) %>%
  rename(rmse = error.power) %>%
  #gather(model, rmse, error.exp, error.power) %>%
  select(metric, case, sizetype, rmse) %>%
  arrange(rmse) %>%
  mutate(
    case = factor(case, levels = c("global", "same", "different")),
    sizetype = factor(sizetype, levels = c("all", "mobile", "freq"))
  ) %>%
  arrange(case, sizetype, rmse) %>%
  group_by(case, sizetype) %>%
  mutate(
    rmse = ifelse(min(rmse) == rmse, paste0("*", round(rmse, 3)), as.character(round(rmse, 3)))
  ) %>%
  group_by(metric, case, sizetype) %>%
  ungroup() %>%
  pivot_wider(names_from = c("case", "sizetype"), values_from = "rmse") %>%
  #select(-id) %>%
  mutate(
    metric = factor(metric, metric_levels, metric_labels)
  ) %>%
  rename(" " = metric)

rmse.hux <- as_hux(rmse.tab) %>%
  set_contents(1, 2:10, rep(c("Global", "Domestic", "International"), 3)) %>%
  insert_row("", "All", "", "", "Mobile only", "", "", "Raw freq", "", "", after = 0) %>% 
  merge_cells(1, 2:4) %>% 
  merge_cells(1, 5:7) %>% 
  merge_cells(1, 8:10) %>%
  set_tb_padding(1, everywhere, 0) %>% 
  set_bold(1, everywhere) %>%
  map_align(by_cols(c("left", rep("right", 9)))) %>%
  set_align(1, everywhere, "center") %>% 
  #map_align(row = 1, by_cols(rep("center", 10))) %>%
  set_header_rows(1:2, TRUE) %>% 
  set_header_cols(1, TRUE) %>% 
  style_headers(bold = TRUE) %>%
  set_right_border(1:12, 1, value = 0.5) %>%
  set_right_border(1:12, 4, value = 0.5) %>%
  set_right_border(1:12, 7, value = 0.5) %>%
  set_bottom_border(2, 1:10, value = 0.5)

rmse.hux %>%
  print_latex()
```