### 2021-09-01 notebooks/embedding_boomerang_plots.rmd
update notebooks

---
title: "Mobility Boomerang Plots"
output: html_notebook
---

```{r}
library(tidyverse)
library(readr)

#df <- read_csv("/Users/dakotamurray/Desktop/l2-norm_vs_size.csv")

meta <- read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/institution_lookup_with_states.txt", delim = "\t")

df <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Derived/Embeddings/precedence/factors/d300_ws1_precedence_gamma1.0_w2vfactors.csv") %>%
  select(cwts_org_no, l2norm) %>%
  left_join(meta, by = "cwts_org_no")

types <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/org_types.csv")

sizes <- read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/2008-2019_inst_sizes_all.txt", delim = "\t")

df <- df %>% left_join(sizes, by = "cwts_org_no")
country <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/country_metadata.csv")
leiden <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/leiden_ranking.csv")
times <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/times_ranking.csv")
head(df)

```

```{r}

labels <- c("Harvard University", "University of Alaska, Fairbanks", "Reykjavik University",
            "University of Oxford", "Princeton University", "Shanghai Jiao Tong University", "Novosibirsk State University", "University of Louisville",  "University of Tennessee, Knoxville", "Barry University", "Appalachian State University", "University of Michigan-Flint", "New York Medical College")

plotdata <- df %>%
  filter(country_iso_alpha == "USA") %>%
  filter(org_type == "University") %>%
  #left_join(meta, by = c("name" = "full_name")) %>%
  #left_join(leiden, by = c("cwts_org_no"))
  left_join(times, by = c("cwts_org_no")) %>%
  arrange(desc(total_score)) %>%
  mutate(
    rank = ifelse(!is.na(total_score), row_number(), NA)
  )

set.seed(117)
plotlabs <- plotdata %>%
  filter(full_name %in% labels) %>%
  mutate(
    label = gsub('University', 'Univ', full_name),
    label = gsub('(.{1,24})(\\s|$)', '\\1\n', label),
    label = trimws(label)
  )

plot <- plotdata %>%
  ggplot(aes(x = size, y = l2norm, color = rank, shape = is.na(rank), label = full_name)) +
  geom_point() +
  scale_x_log10(
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  scale_y_continuous(
    limits = c(0, 7),
    breaks = c(2, 4, 6)
  ) +
  viridis::scale_color_viridis(
    name = "Times Rank", 
    option = "C",
    breaks = c(1, 50, 100)
  ) +
  scale_shape_manual(values = c(16, 1)) +
  #ggrepel::geom_label_repel(
  #  data = plotlabs,
  #  aes(label = label),
  #  color = "black",
  #  force = 50,
  #  min.segment.length = 0.1,
  #  alpha = 0.9,
  #  size = 3.5
  #) +
  guides(shape = F,
         color = guide_colorbar(reverse = T)) +
  theme_minimal() +
  annotate(geom = "label", x = 1.6e4, y = 3.55, label = "Harvard", size = 3.5) +
  annotate(geom = "label", x = 1.6e3, y = 3.5, label = "Princeton", size = 3.5) +
  annotate(geom = "label", x = 8.5e3, y = 5.8, label = "Univ Tennessee\nKnoxville", size = 3.5) +
  annotate(geom = "label", x = 1.8e2, y = 1.55, label = "Univ Alaska\nFairbanks", size = 3.5) +
  annotate(geom = "label", x = 4e2, y = 6.0, label = "NY Medical College", size = 3.5) +
  annotate(geom = "label", x = 1.5e1, y = 4.6, label = "Univ Michigan\nFlint", size = 3.4) +
  annotate(geom = "label", x = 6.8e1, y = 2.48, label = "Barry University", size = 3.5) +
  theme(
    text = element_text(family = "Helvetica", size = 12),
    axis.title.x = element_text(angle = 0, size = 14, face = "bold", vjust = 0.5),
    axis.title.x.top = element_text(angle = 0, size = 14, face = "bold", vjust = 0.5),
    axis.text.x.top = element_blank(),
    axis.title.y = element_text(angle = 90, size = 14, face = "bold"),
    axis.title.y.right = element_text(angle = 90, size = 14, face = "bold"),
    axis.text.y.right = element_blank(),
    legend.text = element_text(size = 12),
    legend.title = element_text(face = "bold"),
    legend.position = c(0.11, 0.852),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    axis.line = element_line(color="black", size = 0.3),
    legend.box.background = element_rect(color="black", size=0.1, fill = "white")
  ) +
  annotation_logticks(sides = "b") +
  xlab("# Researchers") +
  ylab("L2 Norm")

plot

ggsave("/Users/dakotamurray/Dropbox/SME-dropbox/Figures/boomerang_ranking_times.pdf", plot, width = 6, height = 6)
```

```{r}
plot <- df %>%
  mutate(
    highlight = ifelse(country_iso_name %in% c("China", "United States"), country_iso_name, "Other"),
    highlight = factor(highlight, levels = c("United States", "China", "Other"))
  ) %>%
  ggplot(aes(x = size, y = l2norm, fill = highlight, shape = highlight, color = highlight, linetype = highlight)) +
  geom_point(alpha = 0.2) +
  stat_smooth(
    method = "loess",
    size = 1.5,
    level = 0.99,
    se = T,
    alpha = 0.6
  ) +
  scale_shape_manual(values = c(16, 16, 1)) +
  scale_color_manual(values = c("#d35400", "#2980b9", "darkgrey")) +
  scale_linetype_manual(values = c(2, 2, 1)) +
  scale_fill_manual(values = c("#d35400", "#2980b9", "grey")) +
  scale_x_log10(
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  scale_y_continuous(
    limits = c(0, 7),
    breaks = c(2, 4, 6)
  ) +
  guides(shape = F) +
  theme_minimal() +
  theme(
    text = element_text(family = "Helvetica", size = 12),
    axis.title.x = element_text(angle = 0, size = 14, face = "bold", vjust = 0.5),
    axis.title.x.top = element_text(angle = 0, size = 14, face = "bold", vjust = 0.5),
    axis.text.x.top = element_blank(),
    axis.title.y = element_text(angle = 90, size = 14, face = "bold"),
    axis.title.y.right = element_text(angle = 90, size = 14, face = "bold"),
    axis.text.y.right = element_blank(),
    legend.text = element_text(size = 12, face = "bold"),
    legend.position = c(0.15, 0.912),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    legend.title = element_blank(),
    axis.line = element_line(color="black", size = 0.3),
    legend.box.background = element_rect(color="black", size=0.1, fill = "white")
  ) +
  annotation_logticks(sides = "b") +
  xlab("# Researchers") +
  ylab("L2 Norm")

plot

ggsave("/Users/dakotamurray/Dropbox/SME-dropbox/Figures/boomerang_global.pdf", plot, width = 6, height = 6)
```

```{r}
ratio <- read_csv("/Users/dakotamurray/Desktop/data_with_ratio.csv")
plot <- ratio %>%
  mutate(ratio = log10(ratio)) %>%
  ggplot(aes(x = ratio, y = `l2-norm`)) +
  geom_hex(
    aes(fill = stat(log10(count)))
  ) +
  # Draw a regression line
  stat_smooth(method = "lm",
              formula = y ~ x,
              color = "darkred",
              alpha = 0.4,
              fill = "darkred",
              se = T,
              level = 0.99,
              size = 0.75,
              fullrange = T)  +
  # Add the r-squared coefficient to the plot
  ggpmisc::stat_poly_eq(formula = y ~ x,
                        geom = "text_npc",
                        aes(label = paste(..rr.label.., sep = "~~~")),
                        parse=TRUE,
                        label.x.npc = 0.95,
                        label.y.npc = 0.96,
                        rr.digits = 2,
                        size = 7
  ) +
  scale_fill_gradientn(colours=c("white", "#7f8c8d"),
                         name = "Freq",
                         breaks = c(0, 1, 2),
                         limits = c(0, 2),
                         labels = function(x) { parse(text=paste0("10^", x)) },
                         na.value=NA
    ) +
  scale_x_continuous(breaks = c(-1.5, 0, 1)) +
  theme_minimal() +
  theme(
    text = element_text(family = "Helvetica", size = 14),
    axis.title.x = element_text(angle = 0, size = 16, face = "bold", vjust = 0.5),
    axis.title.x.top = element_text(angle = 0, size = 16, face = "bold", vjust = 0.5),
    axis.text.x.top = element_blank(),
    axis.title.y = element_text(angle = 90, size = 16, face = "bold"),
    axis.title.y.right = element_text(angle = 90, size = 16, face = "bold"),
    axis.text.y.right = element_blank(),
    legend.text = element_text(size = 14),
    #legend.position = c(0.16, 0.35),
    legend.position = "right",
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_rect(size = 0.5, color = "black", fill = NA),
    legend.title = element_text(face = "bold"),
    legend.box.background = element_rect(color="black", size=0.1, fill = "white")
  ) +
  annotation_logticks(sides = "b") +
  xlab("# connections / # researchers") +
  ylab("L2 Norm")

plot 
ggsave("/Users/dakotamurray/Dropbox/SME-dropbox/Figures/ratio_l2.pdf", plot, width = 4, height = 3)
```

```{r}
selected <- (df %>%
  group_by(country_iso_alpha) %>%
  summarize(total = sum(size)) %>%
  top_n(30, total))$country_iso_alpha

df %>%
  filter(country_iso_alpha %in% selected) %>%
  group_by(country_iso_name) %>%
  summarize(
    mu = mean(l2norm)
  ) %>%
  arrange(mu)

plotdata <- df %>%
  filter(country_iso_name != "NULL") %>%
  filter(country_iso_alpha %in% selected) %>%
  rename(country = country_iso_name) %>%
  mutate(
    country = ifelse(country == "Russian Federation", "Russia", country),
    country = ifelse(country == "Korea, Republic of", "South Korea", country),
    country = ifelse(country == "Taiwan, Province of China", "Taiwan", country),
    country = ifelse(country == "Iran, Islamic Republic of", "Iran", country),
  )

mu <- plotdata %>%
  group_by(country) %>%
  summarize(
    mu = mean(l2norm)
  )

plotdata %>%
  left_join(mu, by = "country") %>%
  ggplot(aes(l2norm, group = country)) +
  geom_histogram(aes(y=(..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +
  geom_vline(aes(xintercept = mu)) +
  geom_text(
    data = mu,
    aes(label = paste0("mu==", round(mu, 1))), 
    parse = T,
    x = 1.5, 
    y = 0.18, 
    size = 3
  ) +
  facet_wrap(~country) +
  scale_y_continuous(breaks = c(0, 0.2)) +
  theme_minimal() +
  theme(
      text = element_text(family = "Helvetica", size = 14),
      axis.title.x = element_text(angle = 0, size = 16, face = "bold", vjust = 0.5),
      axis.title.y = element_text(angle = 90, size = 16, face = "bold"),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_blank(),
      panel.border = element_rect(size = 0.5, color = "black", fill = NA),
  ) +
  xlab("L2 Norm of Organization Vectors by Country") +
  ylab("Proportion")

```

```{r}
tab <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Derived/Embeddings/precedence/norm/d300_ws1_precedence_norm_by_country.csv")

tab <- tab %>%
  filter(country %in% selected) %>%
  mutate(
    country = ifelse(country == "Russian Federation", "Russia", country),
    country = ifelse(country == "Korea, Republic of", "South Korea", country),
    country = ifelse(country == "Taiwan, Province of China", "Taiwan", country),
    country = ifelse(country == "Iran, Islamic Republic of", "Iran", country),
  ) %>%
  arrange((l2norm))

tab
write_csv(tab, "~/Desktop/norm_by_country.csv")
```