### 2021-09-02 notebooks/network_centralities.rmd

---
title: "Network Centralities"
output: html_notebook
---

```{r}

library(tidyverse)
library(readr)

meta <- read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/institution_lookup_with_states.txt", delim = "\t")

l2norm  <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Derived/Embeddings/precedence/factors/d300_ws1_precedence_w2vfactors.csv") %>%
  select(cwts_org_no, l2norm)

sizes <- read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/Data/raw/2008-2019_inst_sizes.txt", delim = "\t") %>%
  group_by(cwts_org_no) %>%
  summarize(
    size = mean(person_count)
  )

types <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/org_types.csv")

semaxis <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Derived/SemAxis/precedence/prestige/times_d300_ws1_n5_precedence_semaxis.csv")
country <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/country_metadata.csv")
leiden <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/leiden_ranking.csv")
times <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/times_ranking.csv")
centralities <- read_csv('/Users/dakotamurray/Dropbox/SME-dropbox/Data/Derived/Network/precedence_net_centralities.csv')
impact <- read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/org_impact_scores.txt", delim = "\t")
```

```{r}

# Impact rankings
# Semaxis rankings
# Sizes 
# Metadata 
# Org type
base_data <- impact %>%
  mutate(mncs = as.numeric(mncs)) %>%
  left_join(meta, by = "cwts_org_no") %>%
  filter(country_iso_alpha == "USA") %>%
  left_join(types, by = "org_type") %>%
  left_join(sizes)

  filter(size > 50) %>%

  arrange(desc(mncs)) %>%
  mutate(impact_rank = row_number()) %>%
  left_join(semaxis, by = "cwts_org_no") %>%
  arrange(desc(sim)) %>%
  mutate(semaxis_rank = row_number())

print(dim(df))

# Correlation test
cor.test( ~ impact_rank + semaxis_rank,
            data = df,
            method = "spearman",
            continuity = FALSE,
            conf.level = 0.95)
```

```{r}

thresholds <- c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140, 150)

all.sectors <- data.table::rbindlist(lapply(c("Institute", "Government", "Teaching"), function(sector) {
  for_sector <- data.table::rbindlist(lapply(thresholds, function(x) {
    df <- base_data %>%
      filter(org_type_simplified == sector) %>%
      filter(size > x) %>%
      arrange(desc(mncs)) %>%
      mutate(impact_rank = row_number()) %>%
      left_join(semaxis, by = "cwts_org_no") %>%
      arrange(desc(sim)) %>%
      mutate(semaxis_rank = row_number())

    if (dim(df)[1] > 10) {
      test <- cor.test( ~ impact_rank + semaxis_rank,
              data = df,
              method = "spearman",
              continuity = FALSE,
              conf.level = 0.95)
      stat = test$estimate[[1]]
    } else {
      stat = NA
    }

    return(data.frame(statistic = stat, threshold = x, sector = sector, n = dim(df)[1]))
  }))

}))

plotdata <- all.sectors %>%
  mutate(statistic = abs(statistic)) %>%
  mutate(
    sector = factor(sector, levels = c("Teaching", "Institute", "Government")),
    sector = factor(sector, 
                    labels = c("Regional & Liberal\nArts Colleges", 
                               "Research Institute",
                               "Government Organization"))
  )

plotdata %>%
  ggplot(aes(x = threshold, y = statistic, color = sector)) +
  geom_point() +
  geom_line() +
  geom_text(data = plotdata %>% filter(threshold %in% c(0, 50, 100)),
            aes(label = paste0("n=", n), x = threshold, y = abs(statistic)),
            vjust = -0.5, hjust = -0.25
  ) +
  scale_x_continuous(limits = c(-10, NA)) +
  facet_wrap(~sector) +
  theme_minimal() +
  theme(
    text = element_text(family = 'Helvetica', size = 12),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.position = "none",
    strip.text = element_text(face = "bold", size = 14),
    plot.title = element_text(face = "bold")
  ) +
  xlab("Threshold for # of yearly authors") +
  ylab(latex2exp::TeX("Spearman's $\\rho")) +
  ggtitle("SemAxis rank vs. MNCS rank by size threshold")

ggsave("~/Desktop/rho_impact_semaxis_rank.png", width = 8, height = 4)
```

```{r}
plot <- centralities %>%
  left_join(meta, by = "cwts_org_no") %>%
  left_join(l2norm, by = c("cwts_org_no")) %>%
  mutate(eigen = log10(eigen)) %>%
  ggplot(aes(x = eigen, y = l2norm)) +
  geom_point() +
  #scale_x_continuous(limits = c(-10, 0)) +
  #theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    text = element_text(family = "Helvetica", size = 14),
    axis.title.x = element_text(angle = 0, size = 16, face = "bold", vjust = 0.5),
    axis.title.y = element_text(angle = 90, size = 16, face = "bold"),
  )

plot

#plotly::ggplotly(plot)
```

```{r}
sizes %>%
  left_join(centralities) %>%
  ggplot(aes(x = (size), y = (eigen))) +
  geom_point() +
  scale_x_log10(
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  scale_y_log10(
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  ggpmisc::stat_poly_eq(formula = y ~ x,
                          geom = "text_npc",
                          aes(label = paste(..rr.label.., sep = "~~~")),
                          parse=TRUE,
                          label.x.npc = 0.1,
                          label.y.npc = 0.96,
                          rr.digits = 2,
                          size = 7
  ) +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    text = element_text(family = "Helvetica", size = 14),
    axis.title.x = element_text(angle = 0, size = 16, face = "bold", vjust = 0.5),
    axis.title.y = element_text(angle = 90, size = 16, face = "bold"),
  ) +
  annotation_logticks(sides = "b") +
  xlab("Organization size") +
  ylab("Eigenvector centrality")

```