### 2021-09-02 notebooks/w2v_factors_plots.rmd

---
title: "Mobility W2V Factors"
output: html_notebook
---

```{r}
library(readr)
library(tidyverse)
factors <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Derived/Embeddings/precedence/factors/d300_ws1_precedence_w2vfactors.csv")

meta <- read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/institution_lookup_with_states.txt", delim = "\t")

mobile <- read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/Data/raw/mobile_researcher_trajectories.txt", delim = "\t")
```

8,445

3,709,869

```{r}
mobile2 <- mobile %>%
  group_by(cluster_id) %>%
  summarize(
    count = n()
  )
```

```{r}
mobile %>%
  group_by(cwts_org_no) %>%
  summarize(
    count = n()
  ) %>%
  arrange(desc(count)) %>%
  mutate(
    index = row_number(),
    ind = factor(index)
  ) %>%
  ggplot(aes(x = index, y = count)) +
  geom_point() +
  scale_y_log10() +
  theme_minimal() +
  xlab("Org Index") +
  ylab("Frequency")
```

```{r}
library(readxl)
carnegie <- read_excel("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/CCIHE2018-PublicData.xlsx", sheet = "Data", trim_ws = T)
uni_crosswalk <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/carnegie_cwts_us_uni_crosswalk.csv")
leiden = read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/leiden_ranking.csv")
census <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/usa_states_to_regions.csv")
inst_sizes = read_delim("/Users/dakotamurray/Dropbox/SME-dropbox/Data/raw/2008-2019_inst_sizes.txt", delim = "\t") %>% 
  group_by(cwts_org_no) %>%
  summarize(
    count = mean(person_count)
  )

org_types <- read_csv("/Users/dakotamurray/Dropbox/SME-dropbox/Data/Additional/org_types.csv")
```

```{r}
meta.all <- meta %>% 
  filter(country_iso_alpha == "USA" & org_type_code == "U") %>%
  left_join(uni_crosswalk, by = "cwts_org_no") %>%
  left_join(org_types, by = "org_type") %>%
  left_join(leiden, by = "cwts_org_no") %>%
  left_join(census, by = "region") %>%
  left_join(inst_sizes, by = "cwts_org_no") %>%
  left_join(carnegie, by = c("cc_id" = "UNITID"))

mobile.usa <- mobile %>%
  inner_join(meta.all['cwts_org_no']) %>%
  group_by(cluster_id) %>%
  mutate(
    traj_len = length(cwts_org_no)
  ) %>%
  group_by(cwts_org_no) %>%
  summarize(
    mean_traj_len = mean(traj_len),
    num_traj = length(unique(cluster_id)),
    num_total = n()
  )

mobile.usa

factors.usa <- factors %>%
  inner_join(meta.all, by = "cwts_org_no") %>%
  left_join(mobile.usa, by = "cwts_org_no") %>%
  mutate(
    research = recode(BASIC2018, `15` = "R1", `16` = "R2", `17` = "R3", .default = "Other"),
    urban = recode(LOCALE, 
                   `11` = "City", `12` = "City", `13` = "City",
                   `21` = "Suburb", `22` = "Suburb", `23` = "Suburb",
                   .default = "Rural"),
    urban = factor(urban, levels = c("Rural", "Suburb", "City"))
  ) %>%
  arrange(desc(impact_frac_mncs)) %>%
  mutate(
    leiden_rank = row_number()
  )

factors.usa
```

```{r}
factors.usa %>%
  group_by(census_division) %>%
  mutate(
    med = median(s_i)
  ) %>%
  ungroup() %>%
  mutate(
    type = factor(census_division),
    type = reorder(type, desc(med))
  ) %>%
  ggplot(aes(x = type, y = s_i)) +
  geom_violin() +
  geom_jitter(width = 0.1, alpha = 0.5) +
  geom_boxplot(width=0.1) +
  scale_y_log10(labels = function(x) { parse(text=paste0("10^", log10(x))) }) +
  theme_minimal() +
  theme(
    text = element_text(family = "Helvetica", size = 12),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.title.x = element_blank()
  ) +
  ylab(bquote(~log[10]~(s['in'])))
```

```{r}
factors.usa %>%
  group_by(urban) %>%
  mutate(
    med = median(s_i)
  ) %>%
  ungroup() %>%
  mutate(
    type = factor(urban),
    type = reorder(type, desc(med))
  ) %>%
  filter(!is.na(type)) %>%
  ggplot(aes(x = type, y = s_i)) +
  geom_violin() +
  geom_jitter(width = 0.1, alpha = 0.5) +
  geom_boxplot(width=0.1) +
  scale_y_log10(labels = function(x) { parse(text=paste0("10^", log10(x))) }) +
  theme_minimal() +
  theme(
    text = element_text(family = "Helvetica", size = 12),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.title.x = element_blank()
  ) +
  ylab(bquote(~log[10]~(s['in'])))
```

```{r}
factors.usa %>%
  group_by(research) %>%
  mutate(
    med = median(s_i)
  ) %>%
  ungroup() %>%
  mutate(
    type = factor(research),
    type = reorder(type, desc(med))
  ) %>%
  filter(!is.na(type)) %>%
  ggplot(aes(x = type, y = s_i)) +
  geom_violin() +
  geom_jitter(width = 0.1, alpha = 0.5) +
  geom_boxplot(width=0.1) +
  scale_y_log10(labels = function(x) { parse(text=paste0("10^", log10(x))) }) +
  theme_minimal() +
  theme(
    text = element_text(family = "Helvetica", size = 12),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks = element_line(),
    axis.ticks.length=unit(-0.25, "cm"), 
    axis.ticks.margin=unit(0.5, "cm")
  ) +
  ylab(bquote(~log[10]~(s['in'])))
```

```{r}
plot <- factors.usa %>%
  gather(key, measure, s_i, gravity_potential) %>%
  mutate(
    #measure = log10(measure),
    key = factor(key, 
                 levels = c("s_i", "gravity_potential"),
                 labels = c("s_i", "phi"))
  ) %>%
  ggplot(aes(x = num_traj, y = measure, label = full_name)) +
  geom_point() +
  facet_wrap(~key) +
  scale_x_log10(
    #breaks = c(1, 5)
  ) +
  scale_y_log10(
    #breaks = c(0, 8)
  ) +
  theme_minimal() +
  theme(
    text = element_text(family = "Helvetica"),
    strip.text = element_text(face = "bold"),
    axis.title = element_text(face = "bold"),
    panel.grid = element_blank(),
    panel.border = element_rect(size = 0.5, fill = NA)
  ) +
  xlab("# Trajectories with University") +
  ylab("Measure")

plot
plotly.plot <- ggplotly(plot, tooltip = "all")
plotly.plot
```

```{r}

plot <- factors.usa %>%
  ggplot(aes(x = s_i, y = mean_traj_len, label = full_name)) +
  geom_point() +
  scale_x_log10() +
  theme_minimal() +
  theme(
    text = element_text(family = "Helvetica"),
    strip.text = element_text(face = "bold"),
    axis.title = element_text(face = "bold"),
    panel.grid = element_blank(),
    panel.border = element_rect(size = 0.5, fill = NA)
  ) +
  ylab("Mean Trajectory Length") +
  xlab("s_i")

plot
plotly.plot <- ggplotly(plot, tooltip = "all")
plotly.plot
```

```{r}
factors.usa %>%
  ggplot(aes(x = impact_frac_mncs, y = mean_traj_len)) +
  geom_point()

```

```{r}
data <- factors.usa %>%
  select(s_i, gravity_potential, count, leiden_rank, FALLENR17, GRFTF17, GRCIP4PR, HUM_RSD, OTHER_RSD, STEM_RSD, SOCSC_RSD, `S&ER&D`, `NONS&ER&D`, full_name) %>%
  mutate(s_i = log10(s_i),
         gravity_potential = log10(gravity_potential),
         count = log10(count)) %>%
  na.omit()

library(plotly)
```

```{r}
count <- 0
breaks_fun <- function(x) {
  count <<- count + 1L
  switch(
    floor(count / 2),
    c(1, 4),
    c(0, 6),
    c(1, 400),
    c(0, 120),
    c(0, 2e+06),
    c(0, 1e+05),
    c(0, 6e+4),
    c(0, 20000),
    c(0, 600),
    c(0, 100),
    c(0, 150),
    c(0, 250)
  )
}

plot <- data %>%
  gather(key, value, count, leiden_rank, FALLENR17, GRFTF17, GRCIP4PR, HUM_RSD, OTHER_RSD, STEM_RSD, SOCSC_RSD, `S&ER&D`, `NONS&ER&D`, gravity_potential) %>%
  mutate(
    key = factor(key, 
                 levels = c("count", "gravity_potential", "leiden_rank", "GRCIP4PR",
                            "S&ER&D", "NONS&ER&D", "FALLENR17", "GRFTF17", 
                            "STEM_RSD", "SOCSC_RSD", "HUM_RSD", "OTHER_RSD"),
                 labels = c("Log10(#authors)", "Log10(Φ)", "Leiden Rank",  "#Doctoral Fields",
                            "S&E $ (1000's)", "Non S&E $ (1000's)", "Total Enrollment", "Graduate Enrollment", 
                            "#STEM PhDs", "#Soc. Sci. PhDs", "#Humanities PhDs", "#Other PhDs"))
  ) %>%
  ggplot(aes(x = value, y = s_i, label = full_name)) +
  geom_point(alpha = 0.8, size = 1) +
  facet_wrap(~key, scale = "free_x", nrow = 3) +
  stat_smooth(method = "loess",
              formula = y ~ x,
              color = "dodgerblue4",
              size = 0.5,
              fullrange = T) +
  stat_smooth(method = "lm",
              formula = y ~ x,
              color = "firebrick4",
              size = 1,
              fullrange = T) +
  ggpmisc::stat_poly_eq(formula = y ~ x,
                        geom = "text_npc",
                        aes(label = paste(..rr.label.., sep = "~~~")),
                        parse=TRUE,
                        label.x.npc = 0.95,
                        rr.digits = 1,
                        size = 3,
                        color = "firebrick4"
  ) +
  scale_x_continuous(
    breaks = breaks_fun,
    expand = c(0.1, 0)
  ) +
  scale_y_continuous(
    breaks = c(0, 4, 8),
    labels = function(x) { parse(text=paste0("10^", x)) },
  ) +
  theme_minimal() +
  theme(
    text = element_text(family = "Helvetica", size = 12),
    axis.title = element_text(size = 14, face = "bold"),
    axis.title.x = element_blank(),
    panel.spacing = unit(0.2, "lines"),
    panel.spacing.x = unit(0.5, "lines"),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_rect(size = 0.5, fill = NA),
    strip.text = element_text(face = "bold"),
    axis.ticks = element_line()
  ) +
  ylab(bquote(~log[10]~(s['in'])))

ggsave("~/Desktop/all_factors_pulling.pdf", plot, height = 6, width = 8)
#plotly.plot <- ggplotly(plot, tooltip = "all")
#plotly.plot
#htmlwidgets::saveWidget(plotly.plot, "~/Desktop/all_factors_pulling.html")
```

```{r}
count <- 0
breaks_fun <- function(x) {
  count <<- count + 1L
  switch(
    count,
    c(1, 4),
    c(1, 7),
    c(1, 400),
    c(0, 120),
    c(0, 2e+06),
    c(0, 1e+05),
    c(0, 6e+4),
    c(0, 20000),
    c(0, 600),
    c(0, 100),
    c(0, 150),
    c(0, 250)
  )
}

plot <- data %>%
  gather(key, value, count, leiden_rank, FALLENR17, GRFTF17, GRCIP4PR, HUM_RSD, OTHER_RSD, STEM_RSD, SOCSC_RSD, `S&ER&D`, `NONS&ER&D`, s_i) %>%
  mutate(
    key = factor(key, 
                 levels = c("count", "s_i", "leiden_rank", "GRCIP4PR",
                            "S&ER&D", "NONS&ER&D", "FALLENR17", "GRFTF17", 
                            "STEM_RSD", "SOCSC_RSD", "HUM_RSD", "OTHER_RSD"),
                 labels = c("Log10(#authors)", "Log10(s)", "Leiden Rank",  "#Doctoral Fields",
                            "S&E $ (1000's)", "Non S&E $ (1000's)", "Total Enrollment", "Graduate Enrollment", 
                            "#STEM PhDs", "#Soc. Sci. PhDs", "#Humanities PhDs", "#Other PhDs"))
  ) %>%
  ggplot(aes(x = value, y = gravity_potential, label = full_name)) +
  geom_point(alpha = 0.8, size = 1) +
  facet_wrap(~key, scale = "free_x", nrow = 3) +
  stat_smooth(method = "loess",
              formula = y ~ x,
              color = "dodgerblue4",
              size = 0.5,
              fullrange = T) +
  stat_smooth(method = "lm",
              formula = y ~ x,
              color = "firebrick4",
              size = 1,
              fullrange = T) +
  ggpmisc::stat_poly_eq(formula = y ~ x,
                        geom = "text_npc",
                        aes(label = paste(..rr.label.., sep = "~~~")),
                        parse=TRUE,
                        label.x.npc = 0.95,
                        rr.digits = 1,
                        size = 3,
                        color = "firebrick4"
  ) +
  scale_x_continuous(
    breaks = breaks_fun,
    expand = c(0.1, 0)
  ) +
  scale_y_continuous(
    breaks = c(0, 4, 8),
    labels = function(x) { parse(text=paste0("10^", x)) },
  ) +
  theme_minimal() +
  theme(
    text = element_text(family = "Helvetica", size = 12),
    axis.title = element_text(size = 14, face = "bold"),
    axis.title.x = element_blank(),
    panel.spacing = unit(0.2, "lines"),
    panel.spacing.x = unit(0.5, "lines"),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_rect(size = 0.5, fill = NA),
    strip.text = element_text(face = "bold"),
    axis.ticks = element_line()
  ) +
  ylab(bquote(~log[10]~(phi)))

ggsave("~/Desktop/all_factors_potential.pdf", plot, height = 6, width = 8)
#plotly.plot <- ggplotly(plot, tooltip = "all")
#plotly.plot
#htmlwidgets::saveWidget(plotly.plot, "~/Desktop/all_factors_potential.html")
```

```{r}
data %>%
  rename(measure = "s_i", to.copmare = "gravity_potential")

data
```

```{r}
summary(lm(log(s_i) ~ log(count) + census_division, data = factors.usa))
```

```{r}
summary(lm(log(s_i) ~ log(count) + research, data = factors.usa))
```

```{r}
summary(lm(log(s_i) ~ log(count) + leiden_rank, data = factors.usa))
```

```{r}
summary(lm(log(s_i) ~ log(count) + urban, data = factors.usa))
```

# Total enrollment
```{r}
summary(lm(log10(s_i) ~ log(count) + FALLENR17, data = factors.usa))
```

# Graduate full-time enrollment
```{r}
summary(lm(log10(s_i) ~ log10(count) + GRFTF17, data = factors.usa))
```

# Number of programs in which a graduate degree is conferred
```{r}
summary(lm(log(s_i) ~ log(count) + GRCIP4PR, data = factors.usa))
```

# Number of doctors degrees conferred
```{r}
summary(lm(log(s_i) ~ log(count) + DOCRSDEG, data = factors.usa))
```

# Number of doctors degrees conferred
```{r}
summary(lm(log(s_i) ~ log(count) + HUM_RSD + OTHER_RSD + STEM_RSD + SOCSC_RSD, data = factors.usa))
```

# R&D Expenditure for STEM (S&ER&D) and Non-STEM (NONS$ER&D) research
```{r}
summary(lm(log(s_i) ~ log(count) + `S&ER&D` + `NONS&ER&D`, data = factors.usa))
```