### 2023-08-24 notebooks/revision_analysis.Rmd
many small changes

---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
lookup_file("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Additional/institution_lookup_fixed.txt", delim="\t")
geo <- read_csv("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Derived/SemAxis/precedence/geography/coasts_d300_ws1_gamma1.0_n20_semaxis.csv")

new <- orgs %>%
  filter(country_iso_alpha == "USA") %>%
  left_join(geo, by = "cwts_org_no") %>%
  select(cwts_org_no, sim, longitude)
```

```{r}
# Sectors may be one of: All, University, Teaching, Institute, or Government"
SECTOR <- "All"

# Load organization meta-info
lookup <- read_delim("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Additional/institution_lookup_with_states.txt", delim = "\t", col_types = readr::cols()) %>%
  filter(country_iso_alpha == "USA") %>% # select only the specified country
  select(cwts_org_no, country_iso_alpha, region, org_type_code, org_type, latitude, longitude)

axis2 <- read_csv("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Derived/SemAxis/precedence/prestige/leiden_d300_ws1_n5_gamma1.0_precedence_semaxis.csv", col_types = readr::cols())

org_types <-  read_csv("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Additional/org_types.csv")

# Load the axis data and filter to only specified countries
sims <- read_csv("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Derived/SemAxis/precedence/geography/coasts_d300_ws1_gamma1.0_n20_semaxis.csv", col_types = readr::cols()) %>%
  inner_join(lookup, by = "cwts_org_no") %>%
  inner_join(axis2, by = "cwts_org_no") %>%
  left_join(org_types, by = "org_type")

# Enforce orientation. Not sure how we can automate this, so we will
# likely have to create separate rules for each kind of axis we choose
# I am defining these rules here based on what we know about the data already, namely
# which regions have more and less elite, or which are more or near the coasts.
cali_avg <- mean(subset(sims, region == "California")$sim.x)
mass_avg <- mean(subset(sims, region == "Massachusetts")$sim.x)
if (cali_avg > mass_avg) {
  sims$sim.x <- -sims$sim.x
}

ny_avg <- mean(subset(sims, region == "New York")$sim.y)
bama_avg <- mean(subset(sims, region == "Alabama")$sim.y)
if (bama_avg > ny_avg) {
  sims$sim.y <- -sims$sim.y
}

if (SECTOR != "All") {
  sims <- sims %>% filter(org_type_simplified == SECTOR)  
}

table(sims$org_type_simplified)

#
cor.test(~ longitude + sim.x,
            data = sims,
            method = "spearman",
            continuity = FALSE,
            conf.level = 0.99)
```

```{r}

leiden <- read_csv("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Additional/leiden_ranking.csv")

# Load organization meta-info
lookup <- read_delim("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Additional/institution_lookup_with_states.txt", delim = "\t", col_types = readr::cols()) %>%
  #filter(country_iso_alpha == "CHN") %>% # select only the specified country
  select(cwts_org_no, country_iso_alpha, region, org_type_code, org_type, latitude, longitude)

res_list <- lapply(
  c("USA","CHN","DEU","GBR","JPN","ITA","KOR","ESP","CAN","AUS"),
  function(x) {
    axis <- read_csv(
      sprintf(
      "/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Derived/SemAxis/OTHER_COUNTRIES/precedence/prestige/%s_leiden_d300_ws1_n5_gamma1.0_precedence_semaxis.csv", 
        x), 
      col_types = readr::cols()
    ) # end read_csv
    sims <- lookup %>%
      inner_join(axis, by = "cwts_org_no") %>%
      inner_join(leiden, by = "cwts_org_no") %>%
      filter(country_iso_alpha == x)

    t <- cor.test(~ impact_frac_mncs + sim,
                data = sims,
                method = "spearman",
                continuity = FALSE,
                conf.level = 0.99)

   return(data.frame(
       Country=x,
       `Count of orgs` = dim(sims)[1], 
       `Spearmans rho` = abs(t$estimate)
      )
    )

  }
)

df = data.table::rbindlist(res_list)

library(stargazer)
stargazer(df, summary=F, rownames=F)

```

Individual-level analysis
- Look at the RoG of 
```{r}
people <- read_delim("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Misc/2008-2019_people_details.tsv", delim="\t") %>%
  select(-first_name, -full_name)

times <- read_csv("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Additional/times_ranking.csv")

# Load organization meta-info
lookup <- read_delim("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Additional/institution_lookup_with_states.txt", delim = "\t", col_types = readr::cols()) %>%
  #filter(country_iso_alpha == "CHN") %>% # select only the specified country
  select(cwts_org_no, country_iso_alpha, org_type_code)

traj <- read_delim("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Raw/precedence_mobility_trajectories.txt", delim="\t") %>%
  left_join(lookup, by = "cwts_org_no") %>%
  group_by(cluster_id) %>%
  filter(first(country_iso_alpha) =="USA" & first(org_type_code) == "U") %>%
  distinct(cluster_id, cwts_org_no, .keep_all = TRUE) %>%
  filter(row_number() <= 2) # get first 2 per group
  filter(length(unique(country_iso_alpha)) == 1) # Limit to those who stay within US

# So what is the analysis exactly?
  # Begin with only authors in the United States who have been affiliated with a uni
  # Mark each author by the prestige of their first university...
    # What about age? Could control for it
    # Also can control for total number of moves, which will effect RoG...
    # Perhaps control for gender?

# What is the question we want to answer?
  # Do we want to use radius of gyration, or is the distance from one to the next 
  # enough? What does it mean to do a typical vs. atypical move?
  # Maybe instead of RoG, we use something even simpler...the embedding distance between the moves...
  # Then, we don't have to introduce any new concepts. I don't think gender should be a aprt of it. We have a whole section about prestige, so we should stick to that narrative.

# Maybe something like, "holding all else constant, what is the largest driver of making a typical vs. atypical move?
# By avoiding the RoG, and just using first-second, we can simplify a lot of things. # Depending on where you start, holding other factors equal, is a person more likely
# to travel between a nearby or a distant university, measured by the embedding space. 
```

```{r}
# We need to investigate whether there is some kind of symmetry in the mobility patterns
# of smaller and larger universities...
library(igraph)
library(tidyverse)
# Load organization meta-info
lookup <- read_delim("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Additional/institution_lookup_with_states.txt", delim = "\t", col_types = readr::cols()) %>%
  select(cwts_org_no, wos_name, country_iso_alpha, org_type_code) %>%
  filter(country_iso_alpha == "USA" & org_type_code == "U")
sizes <- read_delim("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Additional/2008-2019_inst_sizes_all.txt", delim="\t")
edges <- read_csv("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Derived/Network/precedence_flows_edgelist.csv")
```

```{r}
# First, get the total degree...
g <- graph.data.frame(edges, directed=FALSE)
df <- data.frame(degree(g, mode="all"))
df$cwts_org_no <- as.numeric(row.names(df))
colnames(df)[1] <- "degree"

sizes %>%
  left_join(lookup, by = "cwts_org_no") %>%
  filter(country_iso_alpha == "USA") %>%
  inner_join(df, by = "cwts_org_no") %>%
  ggplot(aes(x = size, y = degree)) + 
  geom_point() +
  scale_x_log10()
# What this shows is that as the size of the institution increases, the degree centrality also increases...
```

```{r}
code <- 1179
# UMich Flint
umich <- edges %>%
  filter(Source == code | Target == code)

library(DescTools)

res <- lapply(unique(lookup$cwts_org_no), function(org) {
  sub_df <- edges %>%
    filter(Source == org | Target == org)

  return(data.frame(cwts_org_no = org, gini = Gini(sub_df$weight)))
})

all_gini <- data.table::rbindlist(res)

p <- all_gini %>%
  left_join(lookup, by = "cwts_org_no") %>%
  left_join(sizes, by = "cwts_org_no") %>%
  filter(size >= 10) %>%
  ggplot(aes(x = size, y = gini, label=wos_name)) +
  geom_point() +
  scale_x_log10()
# Calculate gini index

ggplotly(p)
# 
# How do we get at the idea of "different contexts"?
# That is, how do we test the idea that its not just the distribution across topics, but the 
# occurrence in very different "topics", that's a bit more abstract to test...
#
# I guess we could use embedding distance? That is, perhaps smaller colleges are connected with 
# a more disparate collection of other universities??? That is, people from smaller colleges 
# are move likely to take atypical paths, whereas those in more elite colleges are likely to 
# follow more standard paths...perhaps this is something??
#
# We could look at international connectivity of universities, the proportion of their 
# researchers that are internationally mobile?? That might correlate with the curve, and 
# get at the idea of comparing different curves

```

```{r}
dist <- read_csv("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Derived/Descriptive/org/aggregate/precedence/aggregate_org_distances_precedence_d300_ws1_gamma1.0_sizeall.csv")

colnames(dist)
```

```{r}
res <- lapply(unique(lookup$cwts_org_no), function(org) {
  sub_df <- dist %>%
    filter(org1 == org | org2 == org)

  return(data.frame(cwts_org_no = org, gini = Gini(sub_df$emb_distance)))
})

all_gini <- data.table::rbindlist(res)

library(plotly)
p <- all_gini %>%
  left_join(lookup, by = "cwts_org_no") %>%
  left_join(sizes, by = "cwts_org_no") %>%
  filter(size >= 10) %>%
  ggplot(aes(x = size, y = gini, label=wos_name)) +
  geom_point() +
  geom_smooth(method="loess") +
  scale_x_log10()

ggplotly(p)

# The idea is that the L2-norm scales with frequency, until we hit bigger more "universal" terms, 
# at which case the scaling is counterbalanced by appearing in a greater diversity of contexts, which 
# in turn "average out" the vector until its norm is small, because it isn't weighted towards any 
# particular context
#
# So the idea says more about big elite colleges, rather than it does about symmetry. So we don't 
# really need to show symmetry between elite and non-elite, but rather simply that elite colleges 
# appear in a greater variety of contexts...
#
#
# So these results show that yes, more elite universities share their researchers with a wider
# variety of places...but can we talk about this further???
#
# But I still don't think it touches on the idea of compression in the embedding space.

```

---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
lookup_file("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Additional/institution_lookup_fixed.txt", delim="\t")
geo <- read_csv("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Derived/SemAxis/precedence/geography/coasts_d300_ws1_gamma1.0_n20_semaxis.csv")

new <- orgs %>%
  filter(country_iso_alpha == "USA") %>%
  left_join(geo, by = "cwts_org_no") %>%
  select(cwts_org_no, sim, longitude)
```

```{r}
# Sectors may be one of: All, University, Teaching, Institute, or Government"
SECTOR <- "All"

# Load organization meta-info
lookup <- read_delim("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Additional/institution_lookup_with_states.txt", delim = "\t", col_types = readr::cols()) %>%
  filter(country_iso_alpha == "USA") %>% # select only the specified country
  select(cwts_org_no, country_iso_alpha, region, org_type_code, org_type, latitude, longitude)

axis2 <- read_csv("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Derived/SemAxis/precedence/prestige/leiden_d300_ws1_n5_gamma1.0_precedence_semaxis.csv", col_types = readr::cols())

org_types <-  read_csv("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Additional/org_types.csv")

# Load the axis data and filter to only specified countries
sims <- read_csv("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Derived/SemAxis/precedence/geography/coasts_d300_ws1_gamma1.0_n20_semaxis.csv", col_types = readr::cols()) %>%
  inner_join(lookup, by = "cwts_org_no") %>%
  inner_join(axis2, by = "cwts_org_no") %>%
  left_join(org_types, by = "org_type")

# Enforce orientation. Not sure how we can automate this, so we will
# likely have to create separate rules for each kind of axis we choose
# I am defining these rules here based on what we know about the data already, namely
# which regions have more and less elite, or which are more or near the coasts.
cali_avg <- mean(subset(sims, region == "California")$sim.x)
mass_avg <- mean(subset(sims, region == "Massachusetts")$sim.x)
if (cali_avg > mass_avg) {
  sims$sim.x <- -sims$sim.x
}

ny_avg <- mean(subset(sims, region == "New York")$sim.y)
bama_avg <- mean(subset(sims, region == "Alabama")$sim.y)
if (bama_avg > ny_avg) {
  sims$sim.y <- -sims$sim.y
}

if (SECTOR != "All") {
  sims <- sims %>% filter(org_type_simplified == SECTOR)  
}

table(sims$org_type_simplified)

#
cor.test(~ longitude + sim.x,
            data = sims,
            method = "spearman",
            continuity = FALSE,
            conf.level = 0.99)
```

```{r}

leiden <- read_csv("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Additional/leiden_ranking.csv")

# Load organization meta-info
lookup <- read_delim("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Additional/institution_lookup_with_states.txt", delim = "\t", col_types = readr::cols()) %>%
  #filter(country_iso_alpha == "CHN") %>% # select only the specified country
  select(cwts_org_no, country_iso_alpha, region, org_type_code, org_type, latitude, longitude)

res_list <- lapply(
  c("USA","CHN","DEU","GBR","JPN","ITA","KOR","ESP","CAN","AUS"),
  function(x) {
    axis <- read_csv(
      sprintf(
      "/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Derived/SemAxis/OTHER_COUNTRIES/precedence/prestige/%s_leiden_d300_ws1_n5_gamma1.0_precedence_semaxis.csv", 
        x), 
      col_types = readr::cols()
    ) # end read_csv
    sims <- lookup %>%
      inner_join(axis, by = "cwts_org_no") %>%
      inner_join(leiden, by = "cwts_org_no") %>%
      filter(country_iso_alpha == x)

    t <- cor.test(~ impact_frac_mncs + sim,
                data = sims,
                method = "spearman",
                continuity = FALSE,
                conf.level = 0.99)

   return(data.frame(
       Country=x,
       `Count of orgs` = dim(sims)[1], 
       `Spearmans rho` = abs(t$estimate)
      )
    )

  }
)

df = data.table::rbindlist(res_list)

library(stargazer)
stargazer(df, summary=F, rownames=F)

```

Individual-level analysis
- Look at the RoG of 
```{r}
people <- read_delim("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Misc/2008-2019_people_details.tsv", delim="\t") %>%
  select(-first_name, -full_name)

times <- read_csv("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Additional/times_ranking.csv")

# Load organization meta-info
lookup <- read_delim("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Additional/institution_lookup_with_states.txt", delim = "\t", col_types = readr::cols()) %>%
  #filter(country_iso_alpha == "CHN") %>% # select only the specified country
  select(cwts_org_no, country_iso_alpha, org_type_code)

traj <- read_delim("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Raw/precedence_mobility_trajectories.txt", delim="\t") %>%
  left_join(lookup, by = "cwts_org_no") %>%
  group_by(cluster_id) %>%
  filter(first(country_iso_alpha) =="USA" & first(org_type_code) == "U") %>%
  distinct(cluster_id, cwts_org_no, .keep_all = TRUE) %>%
  filter(row_number() <= 2) # get first 2 per group
  filter(length(unique(country_iso_alpha)) == 1) # Limit to those who stay within US

# So what is the analysis exactly?
  # Begin with only authors in the United States who have been affiliated with a uni
  # Mark each author by the prestige of their first university...
    # What about age? Could control for it
    # Also can control for total number of moves, which will effect RoG...
    # Perhaps control for gender?

# What is the question we want to answer?
  # Do we want to use radius of gyration, or is the distance from one to the next 
  # enough? What does it mean to do a typical vs. atypical move?
  # Maybe instead of RoG, we use something even simpler...the embedding distance between the moves...
  # Then, we don't have to introduce any new concepts. I don't think gender should be a aprt of it. We have a whole section about prestige, so we should stick to that narrative. 
  # Perhaps could we do something w/ attrition???

# Maybe something like, "holding all else constant, what is the largest driver of making a typical vs. atypical move?
# By avoiding the RoG, and just using first-second, we can simplify a lot of things. # Depending on where you start, holding other factors equal, is a person more likely
# to travel between a nearby or a distant university, measured by the embedding space. 
```

```{r}
# We need to investigate whether there is some kind of symmetry in the mobility patterns
# of smaller and larger universities...
library(igraph)
library(tidyverse)
# Load organization meta-info
lookup <- read_delim("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Additional/institution_lookup_with_states.txt", delim = "\t", col_types = readr::cols()) %>%
  select(cwts_org_no, wos_name, country_iso_alpha, org_type_code) %>%
  filter(country_iso_alpha == "USA" & org_type_code == "U")
sizes <- read_delim("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Additional/2008-2019_inst_sizes_all.txt", delim="\t")
edges <- read_csv("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Derived/Network/precedence_flows_edgelist.csv")
```

```{r}
# First, get the total degree...
g <- graph.data.frame(edges, directed=FALSE)
df <- data.frame(degree(g, mode="all"))
df$cwts_org_no <- as.numeric(row.names(df))
colnames(df)[1] <- "degree"

sizes %>%
  left_join(lookup, by = "cwts_org_no") %>%
  filter(country_iso_alpha == "USA") %>%
  inner_join(df, by = "cwts_org_no") %>%
  ggplot(aes(x = size, y = degree)) + 
  geom_point() +
  scale_x_log10()
# What this shows is that as the size of the institution increases, the degree centrality also increases...
```

```{r}
library(DescTools)

lookup <- read_delim("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Additional/institution_lookup_with_states.txt", delim = "\t", col_types = readr::cols()) %>%
  select(cwts_org_no, wos_name, country_iso_alpha, org_type_code) %>%
  filter(country_iso_alpha == "USA" & org_type_code == "U")

res <- lapply(unique(lookup$cwts_org_no), function(org) {
  sub_df <- edges %>%
    filter(Source == org | Target == org)

  return(data.frame(cwts_org_no = org, gini = Gini(sub_df$weight)))
})

all_gini <- data.table::rbindlist(res)

p <- all_gini %>%
  left_join(lookup, by = "cwts_org_no") %>%
  left_join(sizes, by = "cwts_org_no") %>%
  filter(size >= 10) %>%
  ggplot(aes(x = size, y = gini, label=wos_name)) +
  geom_point() +
  scale_x_log10()
# Calculate gini index

ggplotly(p)
# 
# How do we get at the idea of "different contexts"?
# That is, how do we test the idea that its not just the distribution across topics, but the 
# occurrence in very different "topics", that's a bit more abstract to test...
#
# I guess we could use embedding distance? That is, perhaps smaller colleges are connected with 
# a more disparate collection of other universities??? That is, people from smaller colleges 
# are move likely to take atypical paths, whereas those in more elite colleges are likely to 
# follow more standard paths...perhaps this is something??
#
# We could look at international connectivity of universities, the proportion of their 
# researchers that are internationally mobile?? That might correlate with the curve, and 
# get at the idea of comparing different curves

```

```{r}
dist <- read_csv("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Derived/Descriptive/org/aggregate/precedence/aggregate_org_distances_precedence_d300_ws1_gamma1.0_sizeall.csv")

colnames(dist)
```

```{r}
library(plotly)

lookup <- read_delim("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Additional/institution_lookup_with_states.txt", delim = "\t", col_types = readr::cols()) %>%
  select(cwts_org_no, wos_name, country_iso_alpha, org_type_code) %>%
  filter((country_iso_alpha %in% c("USA", "CHN", "EGY", "BRA", "AUS")) & org_type_code == "U")

# What fraction of their total mobility is international???
dist <- dist %>%
  mutate(international = org1_country != org2_country)

res <- lapply(unique(lookup$cwts_org_no), function(org) {
  sub_df <- dist %>%
    filter(org1 == org | org2 == org) %>%
    mutate(international = factor(international, levels = c(TRUE, FALSE))) %>%
    complete(international, fill= list(count=0)) %>%
    group_by(international) %>%
    summarize(count = sum(count)) %>%
    ungroup() %>%
    filter(!is.na(international)) %>%
    pivot_wider(names_from = international, values_from = count) %>%
    mutate(cwts_org_no = org)

  return(sub_df)
})

all_gini <- data.table::rbindlist(res)

p <- all_gini %>%
  mutate(international = `TRUE` / (`TRUE` + `FALSE`)) %>%
  left_join(lookup, by = "cwts_org_no") %>%
  left_join(sizes, by = "cwts_org_no") %>%
  filter(size >= 10) %>%
  ggplot(aes(x = size, y = international, label=wos_name, color=country_iso_alpha)) +
  geom_point() +
  geom_smooth(method="loess") +
  scale_x_log10()

ggplotly(p)

# The idea is that the L2-norm scales with frequency, until we hit bigger more "universal" terms, 
# at which case the scaling is counterbalanced by appearing in a greater diversity of contexts, which 
# in turn "average out" the vector until its norm is small, because it isn't weighted towards any 
# particular context
#
# So the idea says more about big elite colleges, rather than it does about symmetry. So we don't 
# really need to show symmetry between elite and non-elite, but rather simply that elite colleges 
# appear in a greater variety of contexts...
#
#
# So these results show that yes, more elite universities share their researchers with a wider
# variety of places...but can we talk about this further???
#
# But I still don't think it touches on the idea of compression in the embedding space.
# Maybe number of different countries??
```

```{r}
library(plotly)

lookup <- read_delim("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Additional/institution_lookup_with_states.txt", delim = "\t", col_types = readr::cols()) %>%
  select(cwts_org_no, wos_name, country_iso_alpha, org_type_code) %>%
  filter((country_iso_alpha %in% c("USA", "CHN", "EGY", "BRA", "GBR")) & org_type_code == "U")

countries <- unique(c(dist$org1_country, dist$org2_country))
# 
res <- lapply(unique(lookup$cwts_org_no), function(org) {
  sub_df <- dist %>%
    filter(org1 == org | org2 == org) %>%
    filter(org1_country != org2_country) %>%
    mutate(
      # make sure destination is always org2
      destination = ifelse(org1 == org, org2_country, org1_country)
    ) %>%
    mutate(destination = factor(destination, levels=countries)) %>%
    complete(destination, fill=list(count=0)) %>%
    # group by destination
    group_by(destination) %>% 
    summarize(
      cwts_org_no = org,
      count = sum(count)
    ) %>%
    ungroup()

  return(data.frame(cwts_org_no = org, gini = Gini(sub_df$count)))
})

res[[3]]
all_gini <- data.table::rbindlist(res)

p <- all_gini %>%
  left_join(lookup, by = "cwts_org_no") %>%
  left_join(sizes, by = "cwts_org_no") %>%
  filter(size >= 10) %>%
  ggplot(aes(x = size, y = gini, label=wos_name, color=country_iso_alpha)) +
  geom_point() +
  geom_smooth(method="loess") +
  scale_x_log10()

p

all_gini
# So we should calculate "more people from a more narrow set of places", or "more people from a broader 
# set of places..."
#
# So if we do a GINI index, but fill in all missing countries with a zero, then that could be a good way of
# showing this...
# I bet we see something even more stark when we do this gini calculation...
```

```{r}
library(igraph)

#' This adds the gadd graph to the main graph, g, and wires all of its vertices
#' to the central vertex of g
attach_to_center <- function(g, gadd) {
  g <- g + gadd + edges(as.vector(rbind(1, gorder(g) + 1:gorder(gadd))))
}

nIter <- 4
nChild <- 6

# The initial graph
g <- make_empty_graph(5, directed = FALSE) + edges(1,2,1,3,1,4,1,5,2,3,3,4,4,5,5,2)

for (j in 1:nIter) {
  g0 <- g
  for (i in 1:nChild) {
    g <- attach_to_center(g, g0)
  }
}

V(g)
```

```{r}
library(igraph)
library(tidyverse)
# Lets create a random graph with the same degree seq
edges <- read_csv("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Derived/Network/precedence_flows_edgelist.csv")
g0 <- graph.data.frame(edges, directed=FALSE)

#sample_degseq(degree_distribution(g0))
my_fit_g0 <- fit_power_law(degree(g0, mode = "all"), implementation = "plfit")

g1 <- sample_fitness_pl(length(V(g0)), length(E(g0)), my_fit_g0$alpha)
#my_fit_g1 <- fit_power_law(degree(g1, mode = "all"), implementation = "plfit")

# for every node, generate a number of random walkers equal to its degree...
#g1 <- sample_pa(10000, m=100, power=6.83, directed=F)
#g1 <- g0
library(igraph)
traj <- sapply(names(V(g1)), function(x1) {
  d <- as.numeric(degree(g1, x1)) # get degree of node

  # sample trajectories starting at the node
  l <- sapply(c(1:d), function(x2) {
    paste(as.numeric(random_walk(g1, start=x1, steps=3)), collapse=" ")
  })

  return(l)
})

write_graph(g1, file="~/Desktop/murray_sample_network.gml")
write.csv(unlist(traj), "~/Desktop/mobility_network_trajectories_pa-model.csv")
```

```{r}
as.numeric(degree(g1, '1292'))
random_walk(g1, start='1292', steps=3)
```

```{r}
names(sapply(V(g1), function(x1) { x1 }))
new_traj
unlist(traj)
(unlist(traj))

edges %>%
  filter(Source == '1' | Target == '1')
```

```{r}
#hist(log10(degree_distribution(g1)))

```

```{r}
library(tidyverse)
df2 <- read_csv("~/Desktop/mobility_network_trajectories_real_final.csv")

sizes <- read_delim("/Users/dmurray/Dropbox (Personal)/SME-dropbox/Data/Additional/2008-2019_inst_sizes_all.txt", delim="\t") %>%
  mutate(all_size = )
df2
df2 %>%
  mutate(token = as.character(token)) %>%
  left_join(sizes, by = c(cwts_org_no=token)) %>%
  ggplot(aes(x = size, y = norm)) +
  geom_point() +
  scale_x_log10()
```